### 2017-09-27 paper/cospectra-paper.tex
Added manuscript

% mnras_template.tex 
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
%\documentclass[fleqn,usenatbib]{mnras}
\documentclass[12pt]{emulateapj}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
%\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage{yfonts}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

%\usepackage{url}
\usepackage{microtype}
%\usepackage{rotating}
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{tabularx}
\usepackage{xspace}

\usepackage{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\nustar}{\project{NuSTAR}\xspace}
\newcommand{\fermi}{\project{Fermi}\xspace}
\newcommand{\rxte}{\project{RXTE}\xspace}
\newcommand{\xmm}{\project{XMM-Newton}\xspace}
\newcommand{\rosat}{\project{ROSAT}\xspace}
\newcommand{\swift}{\project{Swift}\xspace}
\newcommand{\astrosat}{\project{Astrosat}\xspace}
\newcommand{\ixpe}{\project{IXPE}\xspace}
\newcommand{\nicer}{\project{NICER}\xspace}

\newcommand{\given}{\ensuremath{\,|\,}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\counts}{\ensuremath{y}}
\newcommand{\pars}{\ensuremath{\theta}}
\newcommand{\mean}{\ensuremath{\lambda}}
\newcommand{\likelihood}{\ensuremath{{\mathcal L}}}
\newcommand{\Poisson}{\ensuremath{{\mathcal P}}}
\newcommand{\Uniform}{\ensuremath{{\mathcal U}}}

\newcommand{\Normal}{\ensuremath{{\mathcal N}}}
\newcommand{\bg}{\ensuremath{\mathrm{bg}}}
\newcommand{\word}{\ensuremath{\phi}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

\begin{document}

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[On The Statistical Properties of Cospectra]{On The Statistical Properties of Cospectra}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author{D.~Huppenkothen\altaffilmark{1,2} and M.~Bachetti\altaffilmark{3}}

\altaffiltext{1}{Center for Data Science, New York University, 65 5h Avenue, 7th Floor, New York, NY 10003}
\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
\altaffiltext{3}{INAF-Osservatorio Astronomico di Cagliari, via della Scienza 5, I-09047 Selargius (CA), Italy}

\begin{abstract}
In recent years, the cross spectrum has received considerable attention as a means of characterising the variability of astronomical sources as a function of wavelength. While much has been written about the statistics of time and phase lags, the cospectrum has only recently been understood as means of mitigating instrumental effects dependent on temporal frequency in astronomical detectors, as well as a method of characterizing the coherent variability in two  wavelength ranges on different time scales. In this paper, we lay out the statistical foundations of the cospectrum, starting with the simplest case of detecting a periodic signal in the presence of white noise. This case is especially relevant for detecting faint X-ray pulsars in detectors heavily affected by instrumental effects, including \nustar, \astrosat\ and \ixpe. We show that the statistical distributions of both single and averaged cospectra differ considerably from those for standard power spectra. While a single cospectrum follows a Laplace distribution exactly, averaged cospectra are approximated by a Gaussian distribution only for more than $\sim\!\! 30$ averaged segments, dependent on the number of trials. We provide an instructive example of a quasi-periodic oscillation in \nustar\ and show that applying standard power spectral statistics leads to underestimated tail probabilities for period detection.

\end{abstract}

% These dates will be filled out by the publisher
%\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
%\pubyear{2017}

% Don't change these lines

% Abstract of the paper

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
%\begin{keywords}
%X-rays:binaries -- X-rays:individual -- stars:black holes -- methods:data analysis -- methods:statistical
%\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\section{Introduction}

Time series analysis is one of the primary ways to understand the physical properties of astronomical object in our universe, from exoplanets and stars to black holes and Active Galactic Nuclei (AGN). 
Fourier analysis, especially through the periodogram\footnote{We distinguish in this paper between the \textit{power spectrum}, which describes the process at the source generating variable time series, and the \textit{periodogram}, which denotes a realization of said power spectrum, i.e.\ the time series we actually observe.}, has long been used to find periodic and quasi-periodic signals as well as characterize the stationary stochastic processes often present in accreting systems. 
While in principle, the statistics of the periodogram is well understood and characterized in the literature \citep[e.g.][]{vanderklis1989}, the periodogram is often subject to instrumental effects like dead time that change its statistical properties and thus make statistical inference difficult in practice \citep[e.g.][]{Zhang+95}.

In the past years, the field of \textit{spectral timing} has enjoyed significant success by making it possible to combine both temporal and spectral information in a single model. Within this framework, the complex cross spectrum---defined as the Fourier transform of one time series with the complex conjugate of the Fourier transform of a second time series--holds a central position. The cross spectrum is commonly used to compute phase lags, which are easily converted to time lags and are central for understanding e.g.\ reverberation mapping in accreting black holes \citep[see][for a recent review]{uttley2014}.

The real part of the cross spectrum, also named \textit{cospectrum}, has gained less attention, but can be just as useful. It has long been used to study gravity waves in the Earth's atmosphere \citep[e.g.][]{john2016}, models of the Martian atmosphere \citep[e.g.][]{wang2016}, the solar heliosphere \citep[e.g.][]{vigeesh2017}, surface elevation of arctic sea ice \citep[e.g.][]{ardhuin2016} and drifting snow \citep[e.g.][]{paterna2016} as well as surface gravity waves in beaches \citep[e.g.][]{fiedler2015} and eddy heat flux in the Earth's troposphere \citep[e.g.][]{wang2015,zurita-gotor2017}.
Within astronomy, it has recently been recognized as one approach to mitigate instrumental effects by making use of the independent detectors present in some telescopes. Where in standard analysis, light curves of multiple detectors are summed before Fourier transforming the summed light curve, it is possible to instead Fourier-transform the signal of two independent detectors. 
The resulting powers will be less affected by an instrumental effect called dead time (see details in \citealt{Bachetti+15}), which leads to frequency-dependent changes in the mean and variance of the statistical distributions governing the periodogram. This is relevant for current X-ray detectors such as \nustar\ and \astrosat, but will also be relevant to future missions with multiple detectors like \ixpe. This approach has recently been used in \nustar\ studies of millisecond pulsars \citep{ferrigno2017}, Ultraluminous X-ray Sources \citep{bachetti2016} and X-ray binaries \citep{barriere2015,zoghbi2016,ingram2016,huppenkothen2017,stiele2017}.
Similarly, the cospectrum of two time series taken with the same instrument, but in different wavelength ranges, can be used to characterize the coherent variability in both time series as a function of frequency.

While much has been written on the subject of the statistics of cross spectra and time lags \citep[e.g.][]{epitropakis2016}, the derivation of cospectral statistics is notably absent from the astronomy literature, and most publications assume that either the $\chi^2_2$ distribution used for standard periodograms or a Gaussian distribution for averaged spectra is appropriate. That these assumptions are valid has not been shown until now.

In this paper, we lay out the basic statistical distributions for detecting periodic and narrow quasi-periodic signals in the presence of detector noise (e.g.\ photon counting statistics) for both single and averaged cospectra. We show below that unlike for the periodogram, the statistical distribution for a single cospectrum reduces to a Laplace distribution, while the distribution for averaged cospectra is considerably more complex. We also find that for averaged cospectra consisting of more than $\sim$30 averaged individual segments, the assumption of a Gaussian distribution is indeed appropriate for single-trial tail probabilities and reduces computation overhead. However, for averaged cospectra of fewer segments, more stringent significance thresholds or large numbers of trials, the statistical distribution--and hence the derived $p$-values used for period detection--deviate significantly from a Gaussian distribution.

The paper is laid out as follows. In Section \ref{sec:single_cospectrum}, we derive the PDF of a single cospectrum, and show associated simulations and detection thresholds for period detection in Section \ref{sec:detectionthresholds}. Section \ref{sec:averaged_cospectra} extends the derivation to the common case where the cospectra of several time series are averaged. We show that the statistical distribution indeed changes from a Laplace distribution once multiple cospectra are averaged and becomes consecutively more Gaussian as a larger number of cospectra are included in the average in Section \ref{sec:averaged_detthres}. Finally, Section \ref{sec:nustarqpo} presents a real-world example using simulated \nustar\ data of a quasi-periodic oscillation (QPO) as commonly found in accreting neutron star X-ray binaries. We end in Section \ref{sec:discussion} with a short discussion and conclusion. All figures and results are reproducible, and the associated code can be found online\footnote{\url{https://github.com/dhuppenkothen/cospectra-paper}}.
In a second, forthcoming paper, we will treat the considerably more complex case of cospectra where the time series consist of stochastic variability and show how to model the cospectrum in both a Maximum Likelihood and Bayesian framework.

For the reader looking for the statistical distributions of relevance, who may be only casually interested in the mathematical background, we point to Equations \ref{eqn:csdist} and \ref{eqn:averaged_pdf} for the probability density functions (PDFs) for a single cospectrum and averaged cospectrum, respectively, and Equations \ref{eqn:cospectrum_cdf} and \ref{eqn:averaged_cdf} for the cumulative distribution functions (CDFs) in both cases.

\section{The Statistical Distributions of Cospectral Densities}
\label{sec:whitenoise_cospectra}

\subsection{Statistical Distribution for a Single Cospectrum}
\label{sec:single_cospectrum}
Consider two independently distributed, constant, stationary time series $\mathbf{x} = \{x_k\}_{k=1}^N$ and $\mathbf{y} = \{y_k\}_{k=1}^N$ with $N$ data points taken at simultaneous time intervals $\{t_k\}_{k=1}^N$. Assume for simplicity that the measurements $x_k$ and $y_k$ are normally distributed, such that $x_k \sim \mathcal{N}(\overline{x}, w_x)$ and $y_k \sim \mathcal{N}(\overline{y}, w_y)$ with means $\overline{x}, \overline{y}$ and variances $w_x, w_y$. The data points in the time series $\mathbf{x}$ and $\mathbf{y}$ can be expressed in terms of a Fourier series,

\begin{eqnarray}
x_k & = & \frac{1}{N} \sum_{j}{\mathcal{F}_x(j)} \nonumber \\
x_y & = & \frac{1}{N} \sum_{j}{\mathcal{F}_y(j)}
\end{eqnarray}

\noindent where

\begin{eqnarray}
\mathcal{F}_x(j) &= & \frac{1}{2} (A_{xj} - i B_{xj}) e^{-i\left( \frac{2 \pi j t}{T} \right)} \\
\mathcal{F}_y(j) &= & \frac{1}{2} (A_{yj} - i B_{yj}) e^{-i\left( \frac{2 \pi j t}{T} \right)} \, ,
\end{eqnarray}

\noindent where $i = \sqrt{-1}$, and $A_{xj}, A_{yj}$ and $B_{xj}, B_{yj}$ describe the real and imaginary parts of the Fourier amplitudes, respectively (for a pedagogical introduction into Fourier analysis, see \citealt{vanderklis1989}). The complex cross spectrum is then calculated by multiplying the Fourier transform of light curve $\mathbf{x}$ with the complex conjugate of the Fourier transform of light curve $\mathbf{y}$ (\citealt{vaughan1997,nowak1999}, see also \citealt{uttley2014} for a recent review of spectral timing techniques):

\begin{eqnarray}
\mathcal{F}_x(j) \mathcal{F}_y^*(j) & = & \frac{1}{2} (A_{xj} - i B_{xj}) e^{i \frac{2 \pi j t}{T}} \frac{1}{2} (A_{yj} + i B_{yj}) e^{i \frac{-2 \pi j t}{T}}\nonumber \\ 
		     & = & \frac{1}{4} [ (A_{xj}A_{yj} + B_{xj}B_{yj}) + \\\nonumber
		     & &  i (A_{xj}B_{yj} - A_{yj}B_{xj}) ]
\end{eqnarray}

Note that for strictly real-valued time series, as light curves in astronomy always are, $A_j = A_{-j}$ and $B_j = - B_{-j}$, such that

\begin{equation}
\mathcal{F}_x(j) \mathcal{F}_y^*(j) = \frac{1}{2} \left\{ (A_{xj}A_{yj} + B_{xj}B_{yj}) + i (A_{xj}B_{yj} - A_{yj}B_{xj}) \right\} \, .
\end{equation}

\noindent The real part of this equation is the cross-spectral equivalent of the power spectral density, also called the cospectrum:

\begin{equation}
C_j = \frac{1}{2}  (A_{xj}A_{yj} + B_{xj}B_{yj}) \, .
\end{equation}

\noindent For normally distributed light curves (and indeed for uncertainties coming from a wide range of statistical distributions, including the Poisson distribution, if $N$ is large), the real and imaginary amplitude components are distributed as $A_{xj}, B_{xj} \sim \mathcal{N}(0, \sigma_x)$ with $\sigma_x =  \sqrt{\frac{\sum_{k=1}^{N}{x_k}}{2}}$ and  $A_{yj}, B_{yj} \sim \mathcal{N}(0, \sigma_y)$ with $\sigma_y = \sqrt{\frac{\sum_{k=1}^{N}{y_k}}{2}}$. 
For standard periodograms, $A_{xj} = A_{yj}$ and $B_{xj} = B_{yj}$, and the power spectral density reduces to $P_j = \frac{1}{2} (A_j^2 + B_j^2)$, which is well-known to follow a $\chi^2$ distribution with $2$ degrees of freedom.

Because this condition is not fulfilled for cospectra, we need to derive the probability distribution of the sum of the products over \textit{independent} Gaussian random variables. The probability distribution of the product of two random variables\footnote{We continue the following derivation using the amplitudes $A_j$, but the same arguments apply exactly to the imaginary amplitudes $B_j$.} $Z = A_{xj}A_{yj}$ is called the product distribution, defined as

\begin{eqnarray}
p_Z(z) & = & \int_{-\infty}^{+\infty}{p_X(x) p_Y(\frac{z}{x}) \frac{1}{|x|} dx} \\\nonumber
	   & = &  \int_{-\infty}^{+\infty}{\frac{1}{2 \pi \sigma_x \sigma_y} \exp{-\frac{x^2}{2\sigma_x^2}} \exp{-\frac{(z/x)^2}{2\sigma_y^2}} \frac{1}{|x|} dx} \label{eqn:pz} \, .
\end{eqnarray}

\noindent It can be shown \citep{watson1922,wishart1932} that the integral in Equation \ref{eqn:pz} above can be reduced to

\begin{equation}
\label{eqn:bessel}
P_Z(z) = \frac{K_0\left( \frac{|z|}{\sigma_x \sigma_y}\right)}{\pi \sigma_x \sigma_y} \, ,
\end{equation}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{../figs/cs_dist.png}
\caption{Distribution of Leahy-normalized cospectral densities (left) and power spectral densities (right), respectively, for the simulated data. In dark grey, we show fine-grained histograms of the simulated powers. In red we plot the theoretical probability distribution the simulated powers should follow: A Laplace distribution with $\mu=0$ and $\sigma=2$ for the cospectral densities and a $\chi^2$-distribution with $2$ degrees of freedom for the power spectral densities. The simulated powers adhere very closely to the theoretical predictions.}
\label{fig:csdist}
\end{center}
\end{figure*}

\noindent where $K_0(x) = \int_{0}^{+\infty}{\frac{\cos{(xt)}}{\sqrt{t^2 + 1}} dt}$ is the Bessel function of the second kind of order $0$. We can now use this result to derive the probability density function of $C_j$. In particular, we find that both random variables $Z_j = A_{xj} A_{yj}$ and $Q_j = B_{xj} B_{yj}$ follow the Bessel distribution defined in Equation \ref{eqn:bessel}. Our task is therefore to find the PDF of the sum of two Bessel distributions. The PDF of this sum requires the convolution of the PDFs of each individual random variable being summed. This convolution is difficult to calculate directly for the Bessel distribution defined in Equation \ref{eqn:csdist} above. We instead consider the \textit{moment-generating function} of the PDF, generally defined as

\begin{equation}
M_Z(t) := \mathbb{E}[e^{tZ}] \, 
\end{equation}

\noindent for a random variable $Z$. Consider the sum of any two independent random variables, $S = Z + Q$. While the PDF of $S$ can be found via the convolution of the individual PDFs, it is often simpler to consider the moment-generating function, where the convolution reduces to a simple multiplication operation:

\begin{equation}
M_S(t) = M_Z(t) M_Q(t) \, .
\end{equation}

\noindent The moment-generating function of the Bessel distribution in Equation \ref{eqn:bessel} above is, in the general case \citep{seijas2012} where the means $\mu_x$ and $\mu_y$ are non-zero and the random variables may have unequal variances $\sigma_x \neq \sigma_y$:

\begin{equation}
M_Z(t) = \frac{\exp{\left( \frac{t\mu_x \mu_y + 0.5 (\mu_y^2 \sigma_x^2 + \mu_x^2 \sigma_y^2) t}{1 - t^2 \sigma_x^2 \sigma_y^2} \right)}}{\sqrt{1 - t^2 \sigma_x^2 \sigma_y^2}}\, ,
\end{equation}
but since $\mu_x = \mu_y = 0$ for Fourier amplitudes, this reduces to
\begin{equation}
M_Z(t) =  \frac{1}{\sqrt{1 - t^2 \sigma_x^2 \sigma_y^2}}  \label{eqn:mgf},
\end{equation}

Thus, the moment-generating function for the sum of $Z_j$ and $Q_j$ becomes

\begin{equation}
M_C(t) = M_Z(t) M_Q(t) = \frac{1}{1 - t^2 \sigma_x^2 \sigma_y^2} \, .
\label{eqn:mgt_c}
\end{equation}

\noindent We note that the Laplace distribution is defined as

\[
p_{\mathrm{Laplace}}(x | \mu, b) = \frac{1}{2b} \exp{\left(-\frac{|x - \mu|}{b} \right)}
\]

\noindent and its moment-generating function as

\[
M_\mathrm{Laplace}(t) = \frac{e^{t\mu}}{1 - b^2 t^2} \, .
\]

Comparing this last equation with Equation \ref{eqn:mgt_c}, we find that that Equation \ref{eqn:mgt_c} is equal to the moment-generating function of the Laplace distribution with $\mu = 0$ and $b = \sigma_x \sigma_y$, and hence the cospectral densities follow a Laplace distribution:

\begin{equation}
p(C_j | 0, \sigma_x\sigma_y) = \frac{1}{\sigma_x \sigma_y} \exp{\left(- \frac{|C_j|}{\sigma_x\sigma_y} \right)} 
\label{eqn:laplace}
\end{equation}

\noindent with

\begin{equation}
\sigma_x =  \sqrt{\frac{\sum_{k=1}^{N}{x_k}}{2}} \;\;\; \mathrm{and} \;\;\; \sigma_y =  \sqrt{\frac{\sum_{k=1}^{N}{y_k}}{2}} \, .
\label{eqn:csdist}
\end{equation}

\subsubsection{Detection Thresholds}
\label{sec:detectionthresholds}

Detection thresholds for cospectra will generally be different from those of classical power spectra, because the Laplace distribution tends to be narrower than the equivalent $\chi^2_2$ distribution for single power spectra. To show how the distributions and the corresponding detection thresholds differ, we simulated simple Poisson-distributed light curves. First, we simulated two light curves with a duration of $10\,\mathrm{s}$ and $10^6$ data points each, corresponding to a time resolution of $10^{-5}\,\mathrm{s}$. The light curves have an identical mean count rate of $10^{6} \, \mathrm{counts/s}$, corresponding to $10$ counts per bin. In order to simulate typical measurement uncertainties, we sampled from a Poisson distribution for each time bin with a rate parameter $\lambda = 10$, corresponding to the average counts per bin.

We then calculated both the cospectrum of the two light curves and the periodogram of only the first light curve for comparison. For simplicity, both spectra were computed in Leahy normalization \citep{leahy1983}, which is typically used when searching for (quasi-)periodic signals in time series.
In order to normalize the cospectrum correctly, we used $2/\sqrt{N_{\mathrm{ph}, x}N_{\mathrm{ph}, y}}$, where $N_{\mathrm{ph}, x}$ and $N_{\mathrm{ph}, y}$ are the number of photons of light curves $\mathbf{x}$ and $\mathbf{y}$, respectively, as prescribed by \citet{Bachetti+15}.
In this normalization the powers are distributed as $\chi^2_2$ exactly for the periodogram, and following a Laplace distribution with $\mu=0$ and $\sigma = 1$ for the cospectrum. In Figure \ref{fig:csdist}, we plot the resulting distribution of powers. While the periodogram is only defined for positive values, the Laplace distribution is symmetric around zero, and in general the cospectrum will comprise both positive and negative powers. It is also immediately visible from Figure \ref{fig:csdist} that the probability of obtaining a certain (positive) noise power is always lower for the Laplace distribution than for the $\chi^2$ distribution. In practice, this implies that using the latter where the former is appropriate, we may miss significant periodic signals, because we assume them to be weaker than they are in reality. To demonstrate this, we plot the survival function in Figure \ref{fig:survival}. The survival function, defined in terms of the CDF as $SF(x) = 1 - CDF(x)$, encodes the tail probability of seeing at least a value $x \geq X$. This tail probability is often considered to be the $p$-value of rejecting the null hypothesis that a certain candidate for a periodic signal could be reasonably produced by the noise powers. The CDF for the Laplace distribution with $\mu=0$ is defined as

\begin{equation}
\label{eqn:cospectrum_cdf}
F_{C_j}(x)) = 
  \begin{cases} 
   \frac{1}{2} \exp{\left(\frac{C_j}{\sigma_x \sigma_y}\right)}  & \text{if } C_j < 0 \\
    1 - \frac{1}{2} \exp{-\left( \frac{C_j}{\sigma_x \sigma_y} \right)}   & \text{if } C_j \geq 0
  \end{cases}
\end{equation}

\noindent Much like the PDF, the tail probability is always smaller for the Laplace distribution, indicating that for a given candidate signal, the $p$-value for rejecting the null hypothesis will be stronger than for $\chi^2$-distributed variables. To reinforce this statement, we again simulated two light curves, each again with a duration of $10\,\mathrm{s}$, but this time with only $1000$ data points for simplicity and speed, and a time resolution of $0.01\,\mathrm{s}$. For this simulation, we assumed a mean count rate of $1000\,\mathrm{counts/s}$ or $10$ counts per bin, and additionally introduced a sinusoidal signal with a period of $0.1\,\mathrm{s}$ and a fractional rms amplitude of $a_\mathrm{frac} = 0.055$. Again, this template was used to produce two Poisson-distributed light curves with a rate parameter equal to the number of counts in each bin as defined by the flat continuum and the periodic signal. In Figure \ref{fig:cs_sim}, we show the cospectral densities along with trial-corrected $0.99$ detection thresholds for both the Laplace and $\chi^2$ distribution. If the powers are assumed to follow a $\chi^2$ distribution, as for the periodogram, the candidate at $10 \,\mathrm{Hz}$ would be discounted at the $99\%$ detection threshold, whereas correctly applying the Laplace distribution yields a correct rejection of the null hypothesis at the same detection threshold.

\begin{figure}
\begin{center}
\includegraphics[width=9cm]{../figs/tailprob.png}
\caption{Tail probabilities for the Laplace and $\chi^2$ distributions, respectively. The tail probability, or survival function, is defined as $SF(x) = 1 - CDF(x)$. The tail probability measures the probability of observing a value $x\geq X$, and is often used for detecting periods in power spectra. For the power spectral densities, we plot both the theoretical prediction for the survival function based on the $\chi^2$ distribution (black dashed line), as well as the corrected distribution for power spectra derived in \citet{groth1975} (red solid line). For illustrative purposes, we show a single-trial $95\%$ detection threshold for the Laplace distribution (black solid vertical line) and the $\chi^2$ distribution (black dashed vertical line).}
\label{fig:survival}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=9cm]{../figs/cs_detec.png}
\caption{Cospectrum of two simulated light curves, each with a constant continuum flux of $10$ counts per bin and a periodic signal at $10\,\mathrm{Hz}$. The latter is clearly visible in the cospectrum. We also show the 99\% detection threshold, corrected for a number of trials equal to the number of spectral bins, assuming Laplace-distributed data (red solid line) and $\chi^2$-distributed data (red dashed line). When the latter distribution is assumed, the periodic signal would not be considered a significant detection, because the $\chi^2$ distribution produces a wider distribution of powers. Applying the correct Laplace distribution, however, allows for the detection of weaker signals.}
\label{fig:cs_sim}
\end{center}
\end{figure}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{../figs/avg_dist.png}
\caption{Histogram of cospectral powers (grey) and the theoretical expectation (red) for three different cases. In the right panel, we show the distribution of powers for a single cospectrum, with its expected Laplace distribution from Equation \ref{eqn:laplace}, in the middle and right panel cospectral powers for averaging 10 (middle) and 100 (right) individual cospectra together. In the middle panel, the theoretical expectation of the sampling distribution is given by Equation \ref{eqn:averaged_pdf}, in the right panel by a Gaussian distribution with $\sigma=\sqrt{2/101}$.}
\label{fig:avg_dist}
\end{center}
\end{figure*}

\subsection{Averaged Cospectra}
\label{sec:averaged_cospectra}

The $\chi^2$ distribution used for periodograms has the simple property that sums of $\chi^2$-distributed variables again follow the same distribution, with a different number of degrees of freedom. The same is not true for the Laplace distribution. For $n$ independent and identically distributed (i.i.d.) random variables distributed following a standard Laplace distribution with a mean of $\mu = 0$ and a width of $b = 1$, the distribution of the sums of these random variables can be derived using the fact that a single Laplace random variable $X$ can be rewritten as the difference of two exponential random variables,

\[
X = Z - Z' \; ,
\]
\noindent and thus for $n$ summed random Laplace random variables,

\begin{equation}
T = \sum_{i=1}^{n} X_i = \sum_{i=1}^{n}Z_i - \sum_{i=1}^{n} Z'_i = G_1 - G_2 \, ,
\end{equation}

\noindent where $G1$ and $G2$ are i.i.d.\ standard gamma random variables with a distribution $g(x) = \frac{1}{\Gamma(\nu)} x^{\nu-1} e^{-x}$ and a shape parameter $\nu = n$. For the full derivation of the density, we refer the reader to \citet{kotz2001} and simply state the end result for the PDF for $n$ averaged standard Laplace random variables, $\overline{X}_n$ (see \citealt{kotz2001}, Equations 2.3.25 and 2.3.26):

\begin{equation}
f_{\overline{X}_n}(x) = \frac{n e^{-|nx|}}{(n-1)! 2^n} \sum_{j=0}^{n-1} \frac{(n-1+j)!}{(n-1-j)! j!} \frac{|nx|^{n-1-j}}{2^j} \; \;\;, x \in R \; .
\label{eqn:averaged_pdf}
\end{equation}

For practical purposes, evaluating this PDF for averaged spectra above $n\sim 85$ is difficult numerically, because the factorials and exponents in the sum become very large and small, respectively. However, as we will show in Section \ref{sec:averaged_detthres} below, we find that in practice, when $n \gtrsim 30$, detection thresholds derived from Equation \ref{eqn:averaged_pdf} provide only a negligible difference over that derived from a normal distribution $N(0, \sqrt{2/(n+1)}$, depending on the significance threshold required and the number of trials.

In order to derive tail probabilities useful for hypothesis testing, we require the CDF rather than the PDF. In order to correctly account for the absolute values in the PDF, we split the CDF into two parts: a case where $x < 0$ and a case where $x \geq 0$. The integral $F_{\overline{X}_n}(x) = P ( X \leq x) = \int_{\infty}^{x} f_{\overline{X}_n}(t) dt$ then becomes:

\begin{equation}
\label{eqn:averaged_cdf}
F_{\overline{X}_n}(x)) = 
  \begin{cases} 
   \sum_{j=0}^{n-1} D \frac{1}{n}(2\Gamma(-j+n) - \gamma(-j+n, nx)) \;, \;x \geq 0 \\
     \sum_{j=0}^{n-1} D  \frac{1}{n}\gamma(-j+n, -nx) \;,\; x < 0
  \end{cases}
\end{equation}

\noindent where $\Gamma(l)= (l-1)!$ is the gamma function, $\gamma(l+1, m) = l\gamma(l,m) - l^m e^{-m}$ is the incomplete upper gamma function, and the pre-factor constant $D$ is defined as

\[
D = \frac{n(n-1+j)!}{(n-1-j)! (n-1)! 2^{n+j}} \, .
\]

\noindent As laid out in Section \ref{sec:detectionthresholds}, the tail probability can easily be calculated via the survival function, $\mathrm{SF}(x) = 1 - \mathrm{CDF}(x)$.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{../figs/avg_pvalues.png}
\caption{Predictions for single-trial $p$-values as a function of cospectral power, for the simulated data sets (black solid line) compared to the theoretically predicted $p$-values using the CDF derived in Equation \ref{eqn:averaged_cdf} (red dashed line) as well as the survival function of a Gaussian distribution (black dashed line). The exact distribution matches the empirically derived tail probabilities from simulations within the uncertainties of the simulations. For the case of two averaged cospectra (left panel), a Gaussian approximation is obviously the wrong choice, and will lead to vastly overestimated significances, because the Gaussian PDF drops off much more sharply than the exact distribution. However, as more cospectra are averaged, the resulting distribution becomes more and more similar to that of a Gaussian (middle panel), and for $\sim$30 averaged cospectra, a Gaussian survival function yields a reasonably good approximation to the tail probabilities one would derive from simulations (right panel), up to $p \approx 10^{-4}$.}
\label{fig:avg_pvalue}
\end{center}
\end{figure*}

\subsubsection{Detection Thresholds}
\label{sec:averaged_detthres}

In order to show the way the probability distribution changes as a function of averaged cospectra, we simulate light curves of $10^{5}$ data points and a mean count rate of $100\,\mathrm{counts}/\mathrm{s}$ consisting of pure white noise. We compute $n$ such light curves and average their cospectral powers in order to show the distribution of those powers compared to the expected probability distributions. In Figure \ref{fig:avg_dist}, we show the simulated distribution of Leahy-normalized powers, along with the distributions that describe them. For a single, non-averaged spectrum, we use the Laplace PDF described in Equation \ref{eqn:laplace}. When averaging $10$ cospectra, we use Equation \ref{eqn:averaged_pdf} and show that the theoretical predictions agree with the simulated powers. Finally, for an averaged cospectra consisting of 100 individual light curves, Equation \ref{eqn:averaged_pdf} becomes difficult to compute numerically, and we use a Gaussian PDF instead, which describes the distribution of simulated powers well.

In order to assess the effect on the $p$-values derived from averaged cospectra, we calculate the tail probabilities for the simulated data sets and compare them with the theoretically expected survival function as defined in Equation \ref{eqn:averaged_cdf} as well as a simple Gaussian distribution (Figure \ref{fig:avg_pvalue}). Similar to the results derived by \citet{balakrishnan1986}, we find that for cospectra of more than 30 averaged light curves, a Gaussian distribution yields a reasonably good approximation to the true distribution up to $p \approx 10^{-4}$ with lower overhead. Note, however, that this holds for single-trial probabilities. In general, one will wish to correct for calculating the significance of multiple trial frequencies, requiring the use of more stringent significance threshold. As shown in Figure \ref{fig:avg_pvalue}, the tail probabilities diverge as a function of power, and thus the Gaussian approximation will increasingly overestimate the significance of the signal the higher the threshold is set. Depending on the number of trials used, it is hence advisable to use Equation \ref{eqn:averaged_cdf} as long as it remains numerically stable (depending on implementation, up to $\sim$85 averaged cospectra).

\section{Example: A QPO in \nustar}
\label{sec:nustarqpo}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{../figs/qpo.png}
\caption{Comparison of the averaged periodograms from the two detectors (blue), the periodogram obtained by the sum of the light curves (yellow; both left panel) and the cospectrum (right panel), for $3000\,\mathrm{s}$ of synthetic \nustar data with an incident count rate of $200 \mathrm{counts}\mathrm{s}^{-1}$ and a strong QPO at $200\,\mathrm{Hz}$. 
The QPO has an rms of $\sim$9\% and a high Q factor of $\sim$40. The shape of the periodogram is strongly distorted from the expected flat powerspectrum centred on a constant value of $2$, with deviations of more that $0.5$ in the mean level of the powers common. In dashed lines, we also show the expected (trial-corrected) $99\%$ confidence level for the periodograms (in the same colour, left panel) as well as the cospectrum (red, right panel). The signal is detected in both the periodogram of the combined light curves as well as the cospectrum. However, because the white noise level is variable in the periodogram and the power spectrum dips below the expected level of $2$ around the frequency of the QPO, the signal is not detected as significantly as it should have been if no dead time was present. Assessing the significance is much more straightforward in the cospectrum, which retains a flat baseline.
}
\label{fig:qpo}
\end{center}
\end{figure*}

In order to show the difference of the detection limits with the cospectrum and the power spectrum, we show how a QPO at 200 Hz from a very bright source would appear in \nustar. We simulate a light curve of $T=3000\mathrm{s}$ duration with a time resolution of $\delta t = 0.5 \,\mathrm{ms}$ and an average count rate of $200 \,\mathrm{counts}\,\mathrm{s}^{-1}$. To this constant background we add a quasi-periodic oscillation with a period of $5\,\mathrm{ms}$, a fractional rms amplitude of $f_\mathrm{rms} = 0.15$ and phases randomized using a normal distribution with a mean of $0$ and a width of $\sigma_\mathrm{qpo} = 0.01$. We then simulated photon events using rejection sampling from this light curve using the software package \textit{stingray}\footnote{\url{https://github.com/StingraySoftware/stingray}}, running the function \texttt{stingray.Eventlist.simulate\_times()} twice in order to produce two light curves that are statistically independent, but have the same signal and properties, as we would expect from an instrument with two independent detectors observing the same object. Subsequently we simulated variable deadtime for both light curves with an average time scale of $2.5 \,\mathrm{ms}$ as commonly seen in \nustar\ data \citep{Bachetti+15} using \textit{HENDRICS}\footnote{\url{https://github.com/StingraySoftware/HENDRICS}} \citep{bachetti2015b}. We then produced the periodogram of the summed light curves, the averaged periodogram of the two individual light curves, and the cospectrum of the two light curves. Note that in all three cases we produced averaged periodograms and co-spectra by splitting the light curves into 600 segments of $5\,\mathrm{s}$ length each in order to suppress the variance in the powers and show the effects of dead time more clearly.

The results are shown in Figure~\ref{fig:qpo}. While in all three cases, the QPO is clearly visible, the two periodograms show strong deviations from the expected flat power spectrum. The shape is distorted and requires a precise model of the non-linearly increasing baseline with a non-linearly increasing rms. While in principle, the periodogram of the combined light curves would have a higher significance by a factor of $\sqrt{2}$, the modeling requirements complicate the calculation of the significance of the QPO. 
The baseline of the cospectrum, conversely, is not distorted by dead time, and requires only an estimate of the \textit{local} rms in order to calculate the significance of the QPO using Equation~\ref{eqn:averaged_cdf}: it is sufficient to \textit{multiply the cospectrum around the feature by an estimate of the local standard deviation of the white noise} (which is 1 in the ideal case) to use the equations above with no modifications.

\section{Discussion and Conclusions}
\label{sec:discussion}
We have derived the statistical distribution for the cospectrum, defined as the real part of the cross spectrum. We show that because the Fourier amplitudes being multiplied to derive the cospectrum are now no longer identical (as is the case in the periodogram), the statistical distributions no longer reduce to a simple $\chi^2$ distribution with two degrees of freedom. Instead, we find that the powers in a single cospectrum follow a Laplace distribution with a mean of $\mu=0$ and a width of $\sigma=1$. This has important consequences for period detection. Most importantly, the Laplace distribution is considerably narrower than the $\chi^2_2$ distribution expected for periodograms, and thus the significance of a candidate periodic signal will generally be underestimated when using the latter. Using the correct distribution therefore helps correctly identifying weak signals, which the $\chi^2_2$ distribution would ignore as false negatives. 
Similarly, we find that the sums of Laplace distributions do not follow a similarly simple expression as in the periodogram case, but is considerably more computationally expensive, and may be difficult to estimate numerically when the number $n$ of averaged light curves in the final cospectrum is large. However, we find that for $n \gtrsim 30$, the statistical distribution can be well approximated by a Gaussian distribution, and the resulting tail probabilities used for period detection are very nearly the same as those derived from the exact distribution, up to a tail probability of $p \approx 10^{-4}$. This conclusion, however, depends sensitively on the detection threshold as well as the number of trials: for very small tail probabilities, the two distributions may still deviate significantly. For practical purposes, we suggest using Equation \ref{eqn:averaged_cdf} for at least up to $\sim$30 averaged cospectra, but also for averages of more spectra if significance thresholds for tail probabilities are smaller than $10^{-4}$, or the number of trials is large.

As an example, we have simulated how a QPO would appear in \nustar\ in the presence of dead time, and have shown that the shape of the periodogram is strongly distorted, whereas that of the cospectrum is not (for a longer introduction into the cospectrum and how it can be used in the presence of dead time, see also \citealt{Bachetti+15}). The significance of the QPO is difficult to assess in the periodogram, because of the non-linearities introduced into the power spectrum by the variable dead time. The standard $\chi^2_2$ distributions may either overestimate or underestimate the significance, depending on the shape of the power spectrum at a given frequency, adding complexity to the detection process in the form of finding a non-linear model for the dead time. The cospectrum, on the other hand, only requires an estimate of the local variance in order to use the equations derived above, making it a far more convenient choice for periodicity detection.

The distributions laid out above allow for detecting periodic and narrow quasi-periodic signals in the presence of detector white noise, and especially important in the context of pulsar detection in X-rays, where faint sources may yield marginal detections even in the best of cases. At the same time, as instruments like \astrosat\ and \ixpe\ allow for observations with higher sensitivity, incorporating an accurate treatment of instrumental biases becomes increasingly important, and the cospectral statistics laid out here provide powerful tools to do so.
Notably absent from this discussion, however, is the much more common case where a source exhibits stochastic variability in the form of red noise or notably broadened quasi-periodic oscillations. In this case, the goal is either estimation of the precise properties of the underlying stochastic process, or detection of periods against a background varying stochastically. As has been shown above, the fact that the cross spectrum consists of two different time series complicates the statistical distributions considerably, and this is similarly true cospectra with variability. The exact treatment of this case is beyond the scope of this paper, and will be considered in depth in a forthcoming publication.

\paragraph{Acknowledgements}
The authors thank Thomas Laetsch for helpful suggestions regarding the mathematical derivations.
Daniela Huppenkothen is supported by the James Arthur Postdoctoral Fellowship and the Moore-Sloan Data Science Environment at New York University. 
MB is supported in part by the Italian Space Agency through agreement ASI-INAF n.2017-12-H.0 and ASI-INFN agreement n.2017-13-H.0.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:
\bibliography{cospectra-paper}
\bibliographystyle{apj}

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2017-09-27 paper/cospectra-paper.tex
Fixed variances

@@
 Consider two independently distributed, constant, stationary time series $\mathbf{x} = \{x_k\}_{k=1}^N$ and $\mathbf{y} = \{y_k\}_{k=1}^N$ with $N$ data points taken at simultaneous time intervals $\{t_k\}_{k=1}^N$.
-Assume for simplicity that the measurements $x_k$ and $y_k$ are normally distributed, such that $x_k \sim \mathcal{N}(\overline{x}, w_x)$ and $y_k \sim \mathcal{N}(\overline{y}, w_y)$ with means $\overline{x}, \overline{y}$ and variances $w_x, w_y$.
+Assume for simplicity that the measurements $x_k$ and $y_k$ are normally distributed, such that $x_k \sim \mathcal{N}(\overline{x}, w_x^2)$ and $y_k \sim \mathcal{N}(\overline{y}, w_y^2)$ with means $\overline{x}, \overline{y}$ and variances $w_x^2, w_y^2$.
 The data points in the time series $\mathbf{x}$ and $\mathbf{y}$ can be expressed in terms of a Fourier series,
@@
 
-\noindent For normally distributed light curves (and indeed for uncertainties coming from a wide range of statistical distributions, including the Poisson distribution, if $N$ is large), the real and imaginary amplitude components are distributed as $A_{xj}, B_{xj} \sim \mathcal{N}(0, \sigma_x)$ with $\sigma_x =  \sqrt{\frac{\sum_{k=1}^{N}{x_k}}{2}}$ and  $A_{yj}, B_{yj} \sim \mathcal{N}(0, \sigma_y)$ with $\sigma_y = \sqrt{\frac{\sum_{k=1}^{N}{y_k}}{2}}$.
+\noindent For normally distributed light curves (and indeed for uncertainties coming from a wide range of statistical distributions, including the Poisson distribution, if $N$ is large), the real and imaginary amplitude components are distributed as $A_{xj}, B_{xj} \sim \mathcal{N}(0, \sigma_x^2)$ with $\sigma_x =  \sqrt{\frac{\sum_{k=1}^{N}{x_k}}{2}}$ and  $A_{yj}, B_{yj} \sim \mathcal{N}(0, \sigma_y^2)$ with $\sigma_y = \sqrt{\frac{\sum_{k=1}^{N}{y_k}}{2}}$.
 For standard periodograms, $A_{xj} = A_{yj}$ and $B_{xj} = B_{yj}$, and the power spectral density reduces to $P_j = \frac{1}{2} (A_j^2 + B_j^2)$, which is well-known to follow a $\chi^2$ distribution with $2$ degrees of freedom.

### 2017-11-01 paper/cospectra-paper.tex
Additions to paper in response to referee comments

--- 2017-09-27 paper/cospectra-paper.tex
@@
 It has long been used to study gravity waves in the Earth's atmosphere \citep[e.g.][]{john2016}, models of the Martian atmosphere \citep[e.g.][]{wang2016}, the solar heliosphere \citep[e.g.][]{vigeesh2017}, surface elevation of arctic sea ice \citep[e.g.][]{ardhuin2016} and drifting snow \citep[e.g.][]{paterna2016} as well as surface gravity waves in beaches \citep[e.g.][]{fiedler2015} and eddy heat flux in the Earth's troposphere \citep[e.g.][]{wang2015,zurita-gotor2017}.
+
 Within astronomy, it has recently been recognized as one approach to mitigate instrumental effects by making use of the independent detectors present in some telescopes.
-Where in standard analysis, light curves of multiple detectors are summed before Fourier transforming the summed light curve, it is possible to instead Fourier-transform the signal of two independent detectors.
-The resulting powers will be less affected by an instrumental effect called dead time (see details in \citealt{Bachetti+15}), which leads to frequency-dependent changes in the mean and variance of the statistical distributions governing the periodogram.
-This is relevant for current X-ray detectors such as \nustar\ and \astrosat, but will also be relevant to future missions with multiple detectors like \ixpe.
+Where in standard analysis, light curves of multiple detectors are summed before Fourier transforming the summed light curve, it is possible to instead Fourier-transform the signal of two independent detectors \textbf{within the same instrument observing the same source strictly simultaneously in the same energy band}.
+The resulting \textbf{co-spectral} powers will be less affected by an instrumental effect called \textit{dead time} (see details in \citealt{Bachetti+15}).
+\textbf{In many X-ray detectors, after detection of a photon there is a time interval during which the detector (or in imaging detectors the individual pixel) is blocked from detecting a second photon.
+This time interval is generally called \textit{dead time} (the detector is effectively ``dead'', and incoming photons will not produce a signal) and its characteristic time scale is set by the details of the detector and can range from very short ($2\mu\mathrm{s}$ in the \fermi\ Gamma-Ray Burst Monitor to very long $2.5\mathrm{ms}$ in \nustar).} It leads to frequency-dependent changes in the mean and variance of the statistical distributions governing the periodogram, \textbf{which cannot be mitigated by averaging periodograms of multiple segments}.
+This is relevant for current X-ray missions \textbf{carrying two identical detectors} such as \nustar\ and \astrosat, but will also be relevant to future missions with multiple detectors like \ixpe.
 This approach has recently been used in \nustar\ studies of millisecond pulsars \citep{ferrigno2017}, Ultraluminous X-ray Sources \citep{bachetti2016} and X-ray binaries \citep{barriere2015,zoghbi2016,ingram2016,huppenkothen2017,stiele2017}.
-Similarly, the cospectrum of two time series taken with the same instrument, but in different wavelength ranges, can be used to characterize the coherent variability in both time series as a function of frequency.
+\textbf{It is particularly important for instruments like \nustar\ which have a fairly long dead time and thus exhibit dead time effects the power spectrum at frequencies comparable to those where signals are expected, and at fairly low count rates}.
+Similarly, the cospectrum of two time series taken with the same instrument, but in different wavelength ranges, can be used to characterize the coherent, \textbf{phase-aligned} variability in both time series as a function of frequency.
 
-While much has been written on the subject of the statistics of cross spectra and time lags \citep[e.g.][]{epitropakis2016}, the derivation of cospectral statistics is notably absent from the astronomy literature, and most publications assume that either the $\chi^2_2$ distribution used for standard periodograms or a Gaussian distribution for averaged spectra is appropriate.
+While much has been written on the subject of the statistics of cross spectra and time lags \citep[e.g.][]{epitropakis2016}, the derivation of cospectral statistics is notably absent from the astronomy literature, and most publications \textit{utilizing the approach of forming co-spectra from identical, but independent detectors observing the same source} assume that either the $\chi^2_2$ distribution used for standard periodograms or a Gaussian distribution for averaged spectra is appropriate.
 That these assumptions are valid has not been shown until now.
@@
 However, for averaged cospectra of fewer segments, more stringent significance thresholds or large numbers of trials, the statistical distribution--and hence the derived $p$-values used for period detection--deviate significantly from a Gaussian distribution.
+\textbf{We note that the results below hold only for cases where the signal to be detected is phase-aligned in both light curves.
+This must be true by definition for analyses utilizing data from independent, identical detectors observing the same source simultaneously, but might not strictly be true for observations of the same source e.g.\ in different energy bands.
+We caution the reader that for signals that are not phase-aligned, the resulting power may decrease significantly and result in a non-detectable signal.}
 
@@
 
+\textbf{In the following, we will consider the example of detecting a strictly periodic or very narrow quasi-periodic signal in the presence of simple white noise, as is commonly the case in pulsar searches in X-ray data.
+We assume that both the white noise and the periodic signal are wide-sense stationary and the light curve is evenly sampled.
+While the white noise is also assumed to be strictly stochastic, the signal to be detected may be either deterministic (in the case of a strictly periodic signal) or stochastic (for a quasi-periodic oscillation).
+We then aim to reject a null hypothesis where the observed cospectral power at a given frequency can be explained by white noise alone.}
+
 \subsection{Statistical Distribution for a Single Cospectrum}
 \label{sec:single_cospectrum}
-Consider two independently distributed, constant, stationary time series $\mathbf{x} = \{x_k\}_{k=1}^N$ and $\mathbf{y} = \{y_k\}_{k=1}^N$ with $N$ data points taken at simultaneous time intervals $\{t_k\}_{k=1}^N$.
-Assume for simplicity that the measurements $x_k$ and $y_k$ are normally distributed, such that $x_k \sim \mathcal{N}(\overline{x}, w_x)$ and $y_k \sim \mathcal{N}(\overline{y}, w_y)$ with means $\overline{x}, \overline{y}$ and variances $w_x, w_y$.
-The data points in the time series $\mathbf{x}$ and $\mathbf{y}$ can be expressed in terms of a Fourier series,
+
+Consider two independently distributed, \textbf{evenly-sampled}, constant, stationary time series,
+
+\begin{eqnarray}
+\mathbf{x} &=& \{x_k\}_{k=1}^N \nonumber \\
+\mathbf{y} &=& \{y_k\}_{k=1}^N \nonumber
+\end{eqnarray}
+
+\noindent with $N$ data points taken at simultaneous time intervals $\{t_k\}_{k=1}^N$ \textbf{with a constant time resolution $\Delta t$ and a total duration $T = N\Delta t$}.
+Assume for simplicity that the measurements $x_k$ and $y_k$ are normally distributed, such that
+
+\begin{eqnarray}
+x_k &\sim& \mathcal{N}(\overline{x}, w_x^2)  \nonumber \\
+y_k &\sim& \mathcal{N}(\overline{y}, w_y^2) \nonumber
+\end{eqnarray}
+
+\noindent with means $\overline{x}, \overline{y}$ and variances $w_x^2, w_y^2$. The data points in the time series $\mathbf{x}$ and $\mathbf{y}$ can be expressed in terms of a Fourier series,
 
@@
 \noindent where $i = \sqrt{-1}$, and $A_{xj}, A_{yj}$ and $B_{xj}, B_{yj}$ describe the real and imaginary parts of the Fourier amplitudes, respectively (for a pedagogical introduction into Fourier analysis, see \citealt{vanderklis1989}).
+\textbf{We restrict $\mathcal{F}_x(j)$ and $\mathcal{F}_y(j)$ to frequencies between $\nu_{j=0} = 1/T$ and the Nyquist frequency $\nu_{j=N/2} = 1/(2\Delta t)$.}
 The complex cross spectrum is then calculated by multiplying the Fourier transform of light curve $\mathbf{x}$ with the complex conjugate of the Fourier transform of light curve $\mathbf{y}$ (\citealt{vaughan1997,nowak1999}, see also \citealt{uttley2014} for a recent review of spectral timing techniques):
@@
 
-\noindent For normally distributed light curves (and indeed for uncertainties coming from a wide range of statistical distributions, including the Poisson distribution, if $N$ is large), the real and imaginary amplitude components are distributed as $A_{xj}, B_{xj} \sim \mathcal{N}(0, \sigma_x)$ with $\sigma_x =  \sqrt{\frac{\sum_{k=1}^{N}{x_k}}{2}}$ and  $A_{yj}, B_{yj} \sim \mathcal{N}(0, \sigma_y)$ with $\sigma_y = \sqrt{\frac{\sum_{k=1}^{N}{y_k}}{2}}$.
+\noindent For \textbf{evenly sampled}, normally distributed light curves (and indeed for uncertainties coming from a wide range of statistical distributions, including the Poisson distribution, if $N$ is large), the real and imaginary amplitude components are distributed as $A_{xj}, B_{xj} \sim \mathcal{N}(0, \sigma_x^2)$ with $\sigma_x =  \sqrt{\frac{\sum_{k=1}^{N}{x_k}}{2}}$ and  $A_{yj}, B_{yj} \sim \mathcal{N}(0, \sigma_y^2)$ with $\sigma_y = \sqrt{\frac{\sum_{k=1}^{N}{y_k}}{2}}$.
+\textbf{Note that these distributions for the Fourier amplitudes only hold if the underlying process producing the observations is stochastic and wide-sense stationary (like e.g.\ the white noise case considered here) as is commonly observed in a wide range of astrophysical sources including black hole X-ray binaries and AGN.
+For strictly deterministic processes, the distributions of the Fourier amplitudes will not be centered on $\mu = 0$, and the distributions below will not be correct.}
+
 For standard periodograms, $A_{xj} = A_{yj}$ and $B_{xj} = B_{yj}$, and the power spectral density reduces to $P_j = \frac{1}{2} (A_j^2 + B_j^2)$, which is well-known to follow a $\chi^2$ distribution with $2$ degrees of freedom.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{../figs/cs_dist.png}
+\includegraphics[width=\textwidth]{cs_dist.png}
 \caption{Distribution of Leahy-normalized cospectral densities (left) and power spectral densities (right), respectively, for the simulated data.
@@
 \begin{center}
-\includegraphics[width=9cm]{../figs/tailprob.png}
+\includegraphics[width=9cm]{tailprob.png}
 \caption{Tail probabilities for the Laplace and $\chi^2$ distributions, respectively.
@@
 \begin{center}
-\includegraphics[width=9cm]{../figs/cs_detec.png}
+\includegraphics[width=9cm]{cs_detec.png}
 \caption{Cospectrum of two simulated light curves, each with a constant continuum flux of $10$ counts per bin and a periodic signal at $10\,\mathrm{Hz}$.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{../figs/avg_dist.png}
+\includegraphics[width=\textwidth]{avg_dist.png}
 \caption{Histogram of cospectral powers (grey) and the theoretical expectation (red) for three different cases.
@@
 For practical purposes, evaluating this PDF for averaged spectra above $n\sim 85$ is difficult numerically, because the factorials and exponents in the sum become very large and small, respectively.
-However, as we will show in Section \ref{sec:averaged_detthres} below, we find that in practice, when $n \gtrsim 30$, detection thresholds derived from Equation \ref{eqn:averaged_pdf} provide only a negligible difference over that derived from a normal distribution $N(0, \sqrt{2/(n+1)}$, depending on the significance threshold required and the number of trials.
+However, as we will show in Section \ref{sec:averaged_detthres} below, \textbf{we expect that for large $n$, the Central Limit Theorem implies that the PDF of averaged cospectral powers tends towards a normal distribution.} We find that in practice, when $n \gtrsim 30$, detection thresholds derived from Equation \ref{eqn:averaged_pdf} provide only a negligible difference over that derived from a normal distribution $N(0, \sqrt{2/(n+1)}$, depending on the significance threshold required and the number of trials.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{../figs/avg_pvalues.png}
+\includegraphics[width=\textwidth]{avg_pvalues.png}
 \caption{Predictions for single-trial $p$-values as a function of cospectral power, for the simulated data sets (black solid line) compared to the theoretically predicted $p$-values using the CDF derived in Equation \ref{eqn:averaged_cdf} (red dashed line) as well as the survival function of a Gaussian distribution (black dashed line).
@@
 
+\section{\textbf{Caveats}}
+
+\textbf{All caveats applying more generally to Fourier analysis and to periodograms specifically also apply to cospectra.
+In particular, as for the $\chi^2$ distribution used for period searches in periodograms, the distributions derived here assume that the underlying process is wide-sense stationary and the light curves are simultaneously and evenly sampled.
+These conditions are typically met in most observations taken with X-ray timing instruments like \nustar\ and \rxte\ for problems such as X-ray timing of X-ray binaries, AGN and X-ray pulsar searches.
+Note that this assumes, however, that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers.
+Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}).
+Uneven sampling more generally introduces covariances between adjacent frequencies and renders many of the statistical assumptions underpinning period searches using both in the periodogram and the cospectrum invalid.
+For unevenly sampled data, other methods that do not depend on a regular sampling pattern must be employed (for example the Lomb-Scargle periodogram \citep{lomb1976, scargle1982} or the Bayesian period search in \citet{gregory1992}).}
+
+\textbf{An additional assumption is the phase-alignment of the periodic signal to be detected in both light curves.
+This is always true when the same source is observed simultaneously in the same energy range in independent, but identical detectors on the same spacecraft, as is common practice for instruments like \nustar\, \astrosat\ and \rxte.
+Because the phase of the periodic signal can more generally be energy-dependent, we caution the reader to be careful when producing cospectra of light curves taken for example in different energy ranges.
+Similarly, the phase of a quasi-periodic signal can shift with times, thus observations at different points in time, or of different sources must be analyzed similarly carefully.
+In all these cases, period searches will only be sensitive to signals that are phase-aligned, and will in the worst case be undetectable if the phase-shift is 90 or 270 degrees.
+}
+
 \section{Example: A QPO in \nustar}
@@
 
+In the following, we first simulate a QPO as expected to be observed in a typical \nustar\ observation, and then show similarly real \nustar\ data containing a signal.
+
+\subsection{Simulated Data}
 \begin{figure*}
 \begin{center}
-\includegraphics[width=\textwidth]{../figs/qpo.png}
+\includegraphics[width=\textwidth]{qpo.png}
 \caption{Comparison of the averaged periodograms from the two detectors (blue), the periodogram obtained by the sum of the light curves (yellow; both left panel) and the cospectrum (right panel), for $3000\,\mathrm{s}$ of synthetic \nustar data with an incident count rate of $200 \mathrm{counts}\mathrm{s}^{-1}$ and a strong QPO at $200\,\mathrm{Hz}$.
@@
 The baseline of the cospectrum, conversely, is not distorted by dead time, and requires only an estimate of the \textit{local} rms in order to calculate the significance of the QPO using Equation~\ref{eqn:averaged_cdf}: it is sufficient to \textit{multiply the cospectrum around the feature by an estimate of the local standard deviation of the white noise} (which is 1 in the ideal case) to use the equations above with no modifications.
+
+\subsection{\nustar\ Observations}
+
+[ADD SOME STUFF ABOUT THE OBSERVATIONS.]
 

### 2017-11-01 paper/referee_response.md
draft response to referee

Response to the referee

We are grateful to the referee for their helpful and constructive comments. We believe that some of the issues raised by the referee may be result of a misunderstanding, and have realized that perhaps our text was not as clear as it should have been. We have expanded explanations in several cases, and have added a new section with caveats, as well as a data analysis on real NuSTAR data for illustrative purposes (all changes are marked in bold). Below, we lay out in detail our changes to the manuscript, and respond directly to the points the referee has raised. We hope that this helps solve the misunderstanding and satisfactorily addresses the referee's concern.

> Re. (i) Even for large Signal-to-noise (S/N) the co-spectrum is entirely insensitive to some oscillations with wrong phase. To my knowledge, no else known method suffers
> from such a weakness. Though remedies exist, the distributions discussed in the paper do not apply to them.
and

> 2. Re. (ii) and (iii) Fig. 3 reveals that the simulations were set up so, that the inserted oscillation remains in phase with the reference signal component, resulting with a large positive peak at the corresponding frequency. In the real word, the phases would be randomly distributed and both positive, negative and near null values would be observed.

The primary use case for the distributions defined in the manuscript are cases where observations were taken simulatenously of the same source with independent, identical detectors. This is now standard practice for multiple X-ray missions (including the now defunct RXTE, currently operating NuSTAR and AstroSAT, and future missions like IXPE). The practice of using the cospectrum of time series observed from two independent detectors, rather than the periodogram of the combined time series, was pioneered in NuSTAR as a means to mitigate the effects of dead time (and has since been used is >20 papers). However, the *underlying data stream* from the source is exactly the same. By definition, any periodic signal, be it in an X-ray pulsar or an accreting X-ray binary, must be strictly phase-aligned if it was observed simultaneously with detectors on the same spacecraft. We acknowledge that this might not be true for observations e.g. where the two time series were observed in different energy bands, since periods might be phase-shifted. We have clarified this point in the text at several occasions (in the Introduction, Section 2 as well as the new section "Caveats") and pointed out failure points.

Even in the latter cases, the cases we consider here always consider simultaneous observations of the same object, usually high-energy objects including pulsars, accreting neutron stars, or accreting black holes. While phase shifts are observed for different energy ranges, it is rather unlikely that the phases in the two observed time series are not covariant with each other, that is, while the *phase* itself might vary randomly if the underlying physical process is random, the phase *shift* will be set by specific (deterministic) conditions in the source (e.g. the geometry) and will therefore not be random.

> Re. (ii) The paper deals with a straw man set entirely in virtual reality.
> The quoted references indicate that the authors have access to real observations,
> yet make no attempt to analyze them with the proposed method.

As described above, the common practice of using the cospectrum derived from two identical detectors implies that the case discussed in the paper does, in fact, exist in reality. In particular, the primary reason this manuscript came into being is that we were both approached independently and asked how to define the statistical distributions for the cospectrum in the case of X-ray pulsation searches with NuSTAR. This, combined with the erroneous assumptions made about the statistical distributions used in the current astronomy literature on the topic (primarily the use of the Chi-square and normal distributions fairly indiscriminately), led us to believe that a manuscript laying out the basic statistical distributions to be used is warranted.

In response to the referee's comments, we have analyzed a set of pulsar observations with NuSTAR, and present the results in Section 4.2.

> Re. (iii) The authors make implicit assumptions which neither are obvious nor properly explained in the text. Thus the critical dependence of Co-spectrum on oscillation phase
> in not mentioned nor the fact that the only even-sampled observations are considered.
> This, and points (i) and (ii) are likely to confuse rather then enlighten a reader fresh in the field.

We have attempted to make our assumptions more explicit throughout the text, particularly in Section 2.

> 3. Re. (iii) Editing of Sect. 2.1 till Eq. (4) is so bad that it becomes source of major confusion for the reader. The text mentions a range of observation times without indication that it is implicitly assumed that observations are even-spaced. To demonstrate so suffice a counterexample of N(-) and N(+) of noise observations clustered around +/-T/2. Except for high frequencies, Fourier transform becomes [a exp(-i omega T/2) + b exp(+i omega T/2)] where Var(a)~1/N(-) differs from Var(b)~1/N(+), contrary to statements past Eq. (6).

We have attempted to edit Section 2.1 for clarity, and have explicitly stated that observations must be evenly sampled (and suggest alternative approaches if they are not).

> 4. A subtle flaw, common yet possibly serious of the derivations in the paper stems from assumption of known mean value 0 of x. This holds for simulations yet would be inappropriate for any real observations. Likely the central parts of the distributions would remain largely unaffected, yet their tails, important for the detection criteria, my be seriously affected possibly invalidating the said criteria. While precise accounting for this effect may be difficult, at least some limits to this spoiling effect should be provided.

We believe that there is a misunderstanding here. It is not the time series x that has a mean of zero, which indeed is unphysical, especially for photon counting data. Instead, the *Fourier amplitudes* have a mean of zero at a given frequency. This assumption will be true for any time series that is a realization of a wide-sense stationary stochastic process. For the simplest case of white noise, the Fourier amplitudes will be defined as A_j = \sum(x_k cos(omega_j*t_k)) and B_j = = \sum(x_k sin(omega_j*t_k)) where the sum is over k time intervals and omega_j is an angular frequency. If the x_k data points are distributed according to a normal distribution with mean \mu and variance \sigma^2 (where the mean could vary as a function of time, \mu = \mu_k), the distribution of the data points x_k with the sine or cosine will lead to a new normally distributed variable with a new mean of \mu_k*cos(omega_j*t_k) and a new variance of \mu_k*cos^2(omega_j*t_k). For frequencies between \nu_0 = 1/T up to the Nyquist frequency, this leads to values that are equally distributed around positive and negative means, and the distribution of the sum of all k values will not be exactly zero, but centred around zero. This will be true even when the process is not white noise, but a stationary "red noise" process, because the direction in either positive or negative direction of the time series around a mean is random, and thus the resulting Fourier amplitudes can tend to be either positive or negative.

This assumption underlies all of X-ray timing, and it is worth noting that in neither of us authors has encountered a case where this assumption has not been true, unless the time series was not a random process, not stationary, not evenly sampled or extremely sparse (with <100 photons in total or so). Thus, we conclude that for the vast majority of applications, this assumption will hold. We did add a statement in the relevant Section 2.1 after Equation 6 to make this more explicit.

> 5. while discussing signal from different detectors the authors do not mention an obvious option of adding power spectra to smooth out noise interference.

Averaging periodograms is a standard method to improve the signal-to-noise ratio in period searches. However, in this specific case of dead time, it is ineffective. Dead time occurs when after recording of a photon, it cannot detect another photon within a certain time span (it is, effectively, dead). This leads to a loss of overall photons, but also to a deviation from the typical Poisson process for photon arrival times to a process that involves a characteristic time scale. The resulting periodogram will show an oscillatory pattern over a large range of frequencies (depending on the characteristic time scale of the dead time), and changes both the mean and the variance of the distribution of powers as a function of frequency. Because the effect is always the same at a given frequency, averaging periodograms does not reduce the effect, but rather leads to narrow distributions with an unknown variance around an unknown mean. This is the primary motivation for the cospectrum as an alternative measure unaffected by dead time.

We have added a short description of the effects of dead time in the Introduction for clarification.

> 6. From the discussion in Sect 2.2.1 the reader could not guess that the
> Central Limit Theorem works here and the purpose of the section is mere check when it becomes good enough approximation.

We have added a half-sentence making that explicit.

> 7. While undoubtedly power spectrum is |F(nu)|^2, there exists a broad class of other estimates of PDF as functions of period/frequency. Hence e.g. wikipedia says: 'In signal processing, a periodogramm is an estimate of the spectral density of a signal ...Today, the periodogramm is a component of more sophisticated methods (see spectral estimation)'. Clearly a generic term is needed and periodogramm would fit nicely here.

Throughout the manuscript, we have tried to separate the different Fourier products quite clearly. As explained in Footnote 4 on page 1, we exclusively use the term "power spectrum" for the (stochastic) process generating a time series. Conversely, we exclusively use "periodogram" to denote the square of the absolute value of the Fourier amplitudes derived from a realization of the power spectrum, and say "cospectrum" or "cospectral powers" when we refer to the cospectral equivalent of the periodogram. In line with the signal processing literature, the cospectrum is explicitly not a subset of the periodogram, and we have kept those two concepts separate for clarity, since for understanding the paper it is crucial to make clear when we mean the squared Fourier amplitudes of a single time series, and when we mean the product of the Fourier amplitudes of one time series with the complex conjugate of the Fourier amplitude of a second time series.

### 2017-11-21 paper/cospectra-paper.tex
added two new figures with nustar results, associated notebook and text in paper draft

@@
 % If you need two or more lines of authors, add an extra line using \newauthor
-\author{D.~Huppenkothen\altaffilmark{1,2} and M.~Bachetti\altaffilmark{3}}
+\author{D.~Huppenkothen\altaffilmark{1,2,3} and M.~Bachetti\altaffilmark{4}}
 
@@
 \altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
-\altaffiltext{3}{INAF-Osservatorio Astronomico di Cagliari, via della Scienza 5, I-09047 Selargius (CA), Italy}
+\altaffiltext{3}{DIRAC Institute, Department of Astronomy, University of Washington, 3910 15th Ave NE, Seattle, WA 98195}
+\altaffiltext{4}{INAF-Osservatorio Astronomico di Cagliari, via della Scienza 5, I-09047 Selargius (CA), Italy}
 
@@
 
-\section{\textbf{Caveats}}
-
-\textbf{All caveats applying more generally to Fourier analysis and to periodograms specifically also apply to cospectra.
-In particular, as for the $\chi^2$ distribution used for period searches in periodograms, the distributions derived here assume that the underlying process is wide-sense stationary and the light curves are simultaneously and evenly sampled.
-These conditions are typically met in most observations taken with X-ray timing instruments like \nustar\ and \rxte\ for problems such as X-ray timing of X-ray binaries, AGN and X-ray pulsar searches.
-Note that this assumes, however, that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers.
-Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}).
-Uneven sampling more generally introduces covariances between adjacent frequencies and renders many of the statistical assumptions underpinning period searches using both in the periodogram and the cospectrum invalid.
-For unevenly sampled data, other methods that do not depend on a regular sampling pattern must be employed (for example the Lomb-Scargle periodogram \citep{lomb1976, scargle1982} or the Bayesian period search in \citet{gregory1992}).}
-
-\textbf{An additional assumption is the phase-alignment of the periodic signal to be detected in both light curves.
-This is always true when the same source is observed simultaneously in the same energy range in independent, but identical detectors on the same spacecraft, as is common practice for instruments like \nustar\, \astrosat\ and \rxte.
-Because the phase of the periodic signal can more generally be energy-dependent, we caution the reader to be careful when producing cospectra of light curves taken for example in different energy ranges.
-Similarly, the phase of a quasi-periodic signal can shift with times, thus observations at different points in time, or of different sources must be analyzed similarly carefully.
-In all these cases, period searches will only be sensitive to signals that are phase-aligned, and will in the worst case be undetectable if the phase-shift is 90 or 270 degrees.
-}
-
-\section{Example: A QPO in \nustar}
-\label{sec:nustarqpo}
-
-In the following, we first simulate a QPO as expected to be observed in a typical \nustar\ observation, and then show similarly real \nustar\ data containing a signal.
-
-\subsection{Simulated Data}
 \begin{figure*}
@@
 
+\section{\textbf{Caveats}}
+
+\textbf{All caveats applying more generally to Fourier analysis and to periodograms specifically also apply to cospectra.
+In particular, as for the $\chi^2$ distribution used for period searches in periodograms, the distributions derived here assume that the underlying process is wide-sense stationary and the light curves are simultaneously and evenly sampled.
+These conditions are typically met in most observations taken with X-ray timing instruments like \nustar\ and \rxte\ for problems such as timing of X-ray binaries, AGN and X-ray pulsar searches.
+Note that this assumes, however, that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers.
+Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}).
+Uneven sampling more generally introduces covariances between adjacent frequencies and renders many of the statistical assumptions underpinning period searches using both in the periodogram and the cospectrum invalid.
+For unevenly sampled data, other methods that do not depend on a regular sampling pattern must be employed (for example the Lomb-Scargle periodogram \citep{lomb1976, scargle1982} or the Bayesian period search in \citet{gregory1992}).}
+
+\textbf{An additional assumption is the phase-alignment of the periodic signal to be detected in both light curves.
+This is always true when the same source is observed simultaneously in the same energy range in independent, but identical detectors on the same spacecraft, as is common practice for instruments like \nustar\, \astrosat\ and \rxte.
+Because the phase of the periodic signal can more generally be energy-dependent, we caution the reader to be careful when producing cospectra of light curves taken for example in different energy ranges.
+Similarly, the phase of a quasi-periodic signal can shift with times, thus observations at different points in time, or of different sources must be analyzed similarly carefully.
+In all these cases, period searches will only be sensitive to signals that are phase-aligned, and will in the worst case be undetectable if the phase-shift is 90 or 270 degrees.
+}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{nustar_lc_psd.pdf}
+\caption{\textbf{Upper panel: Light curve of Her X-1 observed with \nustar, observation [XX].
+The observation totals $32.67\,\mathrm{ks}$, the light curve presented here is produced by combining event data from both FPMA and FPMB detectors.
+Gaps in the light curve are caused by the $96.8 \,\mathrm{m}$ orbit around Earth.
+Lower panel: averaged periodogram of $3260$ $10\,\mathrm{s}$-duration segments.
+The bright pulsar produces a very strong periodic signal at $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$ ($P_\mathrm{rot} = 1.24\,\mathrm{s}$) and at the signal's harmonics.
+In order to show the effects of dead time on the underlying detector noise, we have zoomed into the noise region, such that the peaks produced by the pulsar extend beyond the upper edge of the plot.
+At the lowest frequencies, the periodogram deviates from the expected $\chi^2$ distribution centered on $\mu = 2$ because of longer-term, aperiodic variability in the source itself.
+At frequencies above a $\sim$5 Hz or so, dead time strongly modulates the noise powers into an oscillatory pattern, indicating that the use of standard statistical distributions commonly used to find periodic or quasi-periodic signals will lead to biased results.
+}}
+\label{fig:herx1_lc}
+\end{center}
+\end{figure*}
+
+\section{Example: A (Q)PO in \nustar}
+\label{sec:nustarqpo}
+
+In the following, we first simulate a QPO as expected to be observed in a typical \nustar\ observation, and then show similarly real \nustar\ data containing a signal.
+
+\subsection{Simulated Data}
 In order to show the difference of the detection limits with the cospectrum and the power spectrum, we show how a QPO at 200 Hz from a very bright source would appear in \nustar.
@@
 
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{nustar_cospectra.pdf}
+\caption{\textbf{Upper panel: Cospectrum of a $10\,\mathrm{s}$ segment of the \nustar\ observations of Her X-1 (black).
+The cospectrum is created by observing the same source simultaneously, but statistically independently with the two identical FPMA and FPMB detectors.
+The rotational signal of the pulsar appears very clearly at $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$, but unlike the periodogram, the cospectrum is not affected by dead time.
+In red, we show the $p = 10^{-3}$ detection threshold, corrected for $6999$ trials (equivalent to the number of frequencies in the periodogram).
+Lower left:  histogram (grey) of the cospectral densities collected from 3260 individual cospectra between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$.
+In red, we also show the theoretically predicted Laplace distribution.
+Lower right: in black, the histogram of cospectral densities collected from $7$ averaged cospectra, each created from $15$ light curve segments of $200\,\mathrm{s}$ duration.
+Again, only cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ were used in this figure.
+In red, we show the theoretical probability density function defined in Equation \ref{eqn:averaged_pdf} with $n = 15$.
+For both averaged and single cospectra, the observed histogram matches the theoretically expected distributions very closely.
+}}
+\label{fig:herx1_cs}
+\end{center}
+\end{figure*}
+
 \subsection{\nustar\ Observations}
 
-[ADD SOME STUFF ABOUT THE OBSERVATIONS.]
+\begin{bf}
+Hercules X-1 (Her X-1) is a well-studied persistent X-ray binary pulsars \citep{tananbaum1972} in an X-ray binary comprising the neutron star itself and a stellar companion HZ Herculis \citep{davidsen1972,forman1972,bahcall1972} with a mass of $\sim$$2.2 \,\mathrm{M}_\odot$ \citep{reynolds1997,leahy2014} whose type varies between late-type A and early-type B with orbital phase \citep{anderson1994,cheng1995} .
+The neutron star spins with a period of $P = 1.23\,\mathrm{s}$ [REF?] and the system overall exhibits an orbital period of $P_\mathrm{orb} = 1.7\,\mathrm{days}$ \citep{bahcall1972}, along with super-orbital variations on a $\sim$35-day timescale \citep{giacconi1973,scott1999,igna2011}.
+
+Her X-1 was observed with the Nuclear Spectroscopic Telescope Array (\nustar) on XX for XX ks [REF!].
+[ADD DETAILS ABOUT OBSERVATIONS AND DATA ANALYSIS HERE; SAY SOMETHING ABOUT NUSTAR DEAD TIME!].
+Her X-1 was observed for a total of $32.67\,\mathrm{ks}$, counting only good time intervals longer than $10\,\mathrm{s}$.
+In Figure \ref{fig:herx1_lc} (upper panel), we present the light curve of the \nustar\ Her X-1 observations.
+The source varied substantially in brightness during the observations on fairly long time scales (10s to 1000s of seconds), indicating a significant source of red noise at low frequencies.
+
+In the lower panel of Figure \ref{fig:herx1_lc}, we present an averaged periodogram created from a total of $3260$ light curve segments, each $10\,\mathrm{s}$ in duration.
+The periodogram shows very strong peaks at the $1.24\,\mathrm{s}$ rotational period of the pulsar, along with its first five harmonics.
+While at low frequencies, the deviation from the expected noise distribution (a $\chi^2$ distribution centered on $2$) can be explained with the longer-term variability within the source, at frequencies above $5 \,\mathrm{Hz}$ or so, the periodogram displays the typical oscillatory pattern associated with dead time (see also \citealt{Bachetti+15}), suggesting that the standard distributions usually applied to periodograms will not produce unbiased results for these observations.
+
+In Figure \ref{fig:herx1_cs} (upper panel), we present a cospectrum of the Her X-1 \nustar\ observation.
+Because the periodic signal generated by the pulsar dominates the cospectrum, we plot the cospectrum of a short segment of $10\,\mathrm{s}$ duration starting at $\mathrm{MET} = 85740109.54169$ for illustrative purpose to highlight the noise properties.
+Unlike the periodogram, the cospectrum is not affected by dead time, and the statistical distributions defined in Section \ref{sec:whitenoise_cospectra} apply (see also Figure \ref{fig:herx1_cs}, lower left panel and below).
+We also show the trial-corrected detection threshold under the null hypothesis that the cospectrum consists solely of white noise.
+The highest cospectral density of $P = 55.16$ occurs at the frequency of the pulsar, $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$; the probability of observing this power under the null hypothesis is effectively $p \sim 0$ within the limits of numerical accuracy.
+
+In order to test our theoretical predictions for the distributions governing (averaged) cospectra, we first produced $3260$ individual cospectra out of each of the $10\,\mathrm{s}$ segments, and for each cospectrum extracted the cospectral densities in the range $50\,\mathrm{Hz}$ to $700\,\mathrm{Hz}$, where we do not expect significant contamination by the pulsar signal and its harmonics, or by the low-frequency red noise component.
+In order to gain sufficient statistics, we plot a histogram of cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ for all $3260$ segments (a total of $21183500$ powers) in Figure \ref{fig:herx1_cs} (lower left panel), along with the theoretically expected Laplace distribution.
+We find a generally high agreement between the Laplace distribution and the distribution of observed cospectral densities: a Kolmogorov-Smirnov test yields $K= 1.77 \times 10^{-4}$ or a $p$-value of rejecting the null hypothesis that the observed distribution is derived from the theoretically expected Laplace distribution of $p = 0.52$.
+
+In order to repeat the same process for averaged cospectra, we take the $7$ GTIs that are longer than $3000\,\mathrm{s}$  and extract a light curve of exactly $3000\,\mathrm{s}$ from each.
+From each of these light curves, we produce an averaged cospectrum by averaging $15$ consecutive light curve segments of $200\,\mathrm{s}$ duration.
+This yields $7$ averaged cospectra.
+As above, we extract cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ and plot the histogram of cospectral densities extracted from all averaged cospectra (910000 powers) in Figure \ref{fig:herx1_cs} (lower right panel).
+As with the single cospectrum, the histogram for the averaged cospectral densities matches the theoretical prediction very closely.
+\end{bf}
 
@@
 
+\textbf{We show that the same is true for observations of the bright neutron star X-ray binary Her X-1, where the intrinsic brightness of the source leads to strong modulations in the periodogram due to dead time.
+For sources observed with \nustar\ and other instruments with a similar set-up of at least two redundant, identical detectors, one can take advantage of the independent, simultaneous observations in two detectors to form a cospectrum.
+While any (quasi-)periodic signal will appear weaker in the cospectrum than in the periodogram (even in the case, as presented here, where the signals in the two detectors are exactly in phase), it has the substantial advantage that it is unaffected by dead time and other similar detector effects.
+We show that the cospectrum of the observations observed with \nustar\ follow the expected theoretical distributions very closely.}
+
 The distributions laid out above allow for detecting periodic and narrow quasi-periodic signals in the presence of detector white noise, and especially important in the context of pulsar detection in X-rays, where faint sources may yield marginal detections even in the best of cases.
-At the same time, as instruments like \astrosat\ and \ixpe\ allow for observations with higher sensitivity, incorporating an accurate treatment of instrumental biases becomes increasingly important, and the cospectral statistics laid out here provide powerful tools to do so.
+At the same time, as instruments like \astrosat\ and \ixpe, \textbf{with similar set-ups involving multiple identical detectors observing the same source simultaneously,} allow for observations with higher sensitivity, incorporating an accurate treatment of instrumental biases becomes increasingly important, and the cospectral statistics laid out here provide powerful tools to do so.
 Notably absent from this discussion, however, is the much more common case where a source exhibits stochastic variability in the form of red noise or notably broadened quasi-periodic oscillations.
@@
 The authors thank Thomas Laetsch for helpful suggestions regarding the mathematical derivations.
-Daniela Huppenkothen is supported by the James Arthur Postdoctoral Fellowship and the Moore-Sloan Data Science Environment at New York University. 
+DH is supported by the James Arthur Postdoctoral Fellowship and the Moore-Sloan Data Science Environment at New York University. 
+DH acknowledges support from the DIRAC Institute in the Department of Astronomy at the University of Washington.
+The DIRAC Institute is supported through generous gifts from the Charles and Lisa Simonyi Fund for Arts and Sciences, and the Washington Research Foundation.
 MB is supported in part by the Italian Space Agency through agreement ASI-INAF n.2017-12-H.0 and ASI-INFN agreement n.2017-13-H.0.

### 2017-11-21 paper/referee_response.md
Added comments about new NuSTAR section to referee response

@@
 
-The primary use case for the distributions defined in the manuscript are cases where observations were taken simulatenously of the same source with independent, identical detectors.
+The primary use case for the distributions defined in the manuscript are cases where observations were taken simultaneously of the same source with independent, identical detectors.
 This is now standard practice for multiple X-ray missions (including the now defunct RXTE, currently operating NuSTAR and AstroSAT, and future missions like IXPE).
@@
 
-Even in the latter cases, the cases we consider here always consider simultaneous observations of the same object, usually high-energy objects including pulsars, accreting neutron stars, or accreting black holes.
+Even if observations were taken e.g.
+in different energy bands, the cases we consider here always consider simultaneous observations of the same object, usually compact objects including pulsars, accreting neutron stars, or accreting black holes.
 While phase shifts are observed for different energy ranges, it is rather unlikely that the phases in the two observed time series are not covariant with each other, that is, while the *phase* itself might vary randomly if the underlying physical process is random, the phase *shift* will be set by specific (deterministic) conditions in the source (e.g.
 the geometry) and will therefore not be random.
+Phase shifts by exactly 90 or 270 degrees are rare in practice, thus while detectability may be diminished, it is still worthwhile to search for it.
+We have clarified this further in the new Section 3 "Caveats".
 
@@
 
-As described above, the common practice of using the cospectrum derived from two identical detectors implies that the case discussed in the paper does, in fact, exist in reality.
+As described above, the common practice of using the cospectrum derived from two identical detectors implies that we disagree with the referee on the point that the case discussed in the paper does not exist in reality.
 In particular, the primary reason this manuscript came into being is that we were both approached independently and asked how to define the statistical distributions for the cospectrum in the case of X-ray pulsation searches with NuSTAR.
-This, combined with the erroneous assumptions made about the statistical distributions used in the current astronomy literature on the topic (primarily the use of the Chi-square and normal distributions fairly indiscriminately), led us to believe that a manuscript laying out the basic statistical distributions to be used is warranted.
+This, combined with the erroneous assumptions made about the statistical distributions used in the current astronomy literature on the topic (primarily the use of the Chi-square and normal distributions), led us to believe that a manuscript laying out the correct statistical distributions to be used is warranted.
 
-In response to the referee's comments, we have analyzed a set of pulsar observations with NuSTAR, and present the results in Section 4.2.
+In response to the referee's comments, we have analyzed a set of NuSTAR observations of the bright X-ray pulsar Hercules X-1, and present the results in Section 4.2.
+In particular, we find that the pulsar observations are heavily affected by dead time, rendering the standard periodogram largely unusable.
+We also find that the cospectral densities follow the distributions laid out in the paper very closely, indicating that the distributions laid out are useful in practice.
 
@@
 
-We have attempted to make our assumptions more explicit throughout the text, particularly in Section 2.
+We have made our assumptions more explicit throughout the text, particularly in Section 2.
 
@@
 
-We have attempted to edit Section 2.1 for clarity, and have explicitly stated that observations must be evenly sampled (and suggest alternative approaches if they are not).
+We have edited Section 2.1 for clarity, and have explicitly stated that observations must be evenly sampled (and suggest alternative approaches if they are not).
 
@@
 We believe that there is a misunderstanding here.
-It is not the time series x that has a mean of zero, which indeed is unphysical, especially for photon counting data.
+It is not the time series x that has a mean of zero, which indeed is unphysical, especially for photon counting data (and all simulations used in the paper use photon counting statistics and a mean count rate >0).
 Instead, the *Fourier amplitudes* have a mean of zero at a given frequency.
@@
 Because the effect is always the same at a given frequency, averaging periodograms does not reduce the effect, but rather leads to narrow distributions with an unknown variance around an unknown mean.
+As an example, we included the periodogram of the NuSTAR observation of Her X-1, for which we averaged the periodograms for 3260 individual light curve segments.
+The resulting noise distributions are narrow, but deviate very strongly from the expected Chi-square distribution with a mean of 2 and a variance of 2/sqrt(M), where M=3260.
 This is the primary motivation for the cospectrum as an alternative measure unaffected by dead time.
@@
 
-Throughout the manuscript, we have tried to separate the different Fourier products quite clearly.
+Throughout the manuscript, we aim to separate the different Fourier products as clearly as possible.
 As explained in Footnote 4 on page 1, we exclusively use the term "power spectrum" for the (stochastic) process generating a time series.
-Conversely, we exclusively use "periodogram" to denote the square of the absolute value of the Fourier amplitudes derived from a realization of the power spectrum, and say "cospectrum" or "cospectral powers" when we refer to the cospectral equivalent of the periodogram.
+Conversely, we exclusively use "periodogram" to denote the square of the absolute value of the Fourier amplitudes derived from a realization of the power spectrum, and say "cospectrum" or "cospectral densities" when we refer to the cospectral equivalent of the periodogram.
 In line with the signal processing literature, the cospectrum is explicitly not a subset of the periodogram, and we have kept those two concepts separate for clarity, since for understanding the paper it is crucial to make clear when we mean the squared Fourier amplitudes of a single time series, and when we mean the product of the Fourier amplitudes of one time series with the complex conjugate of the Fourier amplitude of a second time series.

### 2017-11-21 paper/cospectra-paper.tex

@@
 
-Her X-1 was observed with the Nuclear Spectroscopic Telescope Array (\nustar) on XX for XX ks [REF!].
-[ADD DETAILS ABOUT OBSERVATIONS AND DATA ANALYSIS HERE; SAY SOMETHING ABOUT NUSTAR DEAD TIME!].
-Her X-1 was observed for a total of $32.67\,\mathrm{ks}$, counting only good time intervals longer than $10\,\mathrm{s}$.
-In Figure \ref{fig:herx1_lc} (upper panel), we present the light curve of the \nustar\ Her X-1 observations.
+Her X-1 was observed by the Nuclear Spectroscopic Telescope Array (\nustar) multiple times in the last years. 
+For this work, we considered the observation taken from UT 2012-09-19 to UT 2012-09-20, one of those used by \citet{fuerst13} to characterize the cyclotron resonance scattering features in the spectrum of the source.
+We downloaded the observation directory for observation ID 30002006002 from the HEASARC and used the FTOOL \texttt{barycorr} on the L2 cleaned science event files (file name ending with \verb|01_cl.evt|) to correct the photon arrival times to the solar system barycenter.
+For our analysis, we considered photons from 3 to 79 keV at most 50\arcsec from the nominal position of the source.
+For this work, we used a total of $32.67\,\mathrm{ks}$ of good time intervals (GTIs), only selecting intervals longer than $10\,\mathrm{s}$.
+In Figure \ref{fig:herx1_lc} (upper panel), we present the light curve of the \nustar Her X-1 observations.
 The source varied substantially in brightness during the observations on fairly long time scales (10s to 1000s of seconds), indicating a significant source of red noise at low frequencies.

### 2017-11-28 paper/referee_response.md
Minor wording changes to referee response

@@
 
-We are grateful to the referee for their helpful and constructive comments.
-We believe that some of the issues raised by the referee may be result of a misunderstanding, and have realized that perhaps our text was not as clear as it should have been.
+We are grateful to the referee for their helpful and constructive comments, which we have addressed below and in the updated version of the manuscript.
+We believe that some of the issues raised by the referee may be result of a misunderstanding, and have realized that our text was not as clear as it should have been.
 We have expanded explanations in several cases, and have added a new section with caveats, as well as a data analysis on real NuSTAR data for illustrative purposes (all changes are marked in bold).
@@
 > from such a weakness. Though remedies exist, the distributions discussed in the paper do not apply to them.
+
 and
@@
 The practice of using the cospectrum of time series observed from two independent detectors, rather than the periodogram of the combined time series, was pioneered in NuSTAR as a means to mitigate the effects of dead time (and has since been used is >20 papers).
-However, the *underlying data stream* from the source is exactly the same.
+However, the *underlying light curve* from the source is exactly the same, and only modified by detectore effects and Poisson counting statistics.
 By definition, any periodic signal, be it in an X-ray pulsar or an accreting X-ray binary, must be strictly phase-aligned if it was observed simultaneously with detectors on the same spacecraft.
@@
 In particular, we find that the pulsar observations are heavily affected by dead time, rendering the standard periodogram largely unusable.
-We also find that the cospectral densities follow the distributions laid out in the paper very closely, indicating that the distributions laid out are useful in practice.
+We also find that the cospectral densities follow the distributions laid out in the paper very closely, indicating that the distributions are useful in practice.
 
@@
 This assumption will be true for any time series that is a realization of a wide-sense stationary stochastic process.
-For the simplest case of white noise, the Fourier amplitudes will be defined as A_j = \sum(x_k cos(omega_j*t_k)) and B_j = = \sum(x_k sin(omega_j*t_k)) where the sum is over k time intervals and omega_j is an angular frequency.
+For the simplest case of white noise, the Fourier amplitudes will be defined as A_j = \sum(x_k cos(omega_j*t_k)) and B_j = \sum(x_k sin(omega_j*t_k)) where the sum is over k time intervals and omega_j is an angular frequency.
 If the x_k data points are distributed according to a normal distribution with mean \mu and variance \sigma^2 (where the mean could vary as a function of time, \mu = \mu_k), the distribution of the data points x_k with the sine or cosine will lead to a new normally distributed variable with a new mean of \mu_k*cos(omega_j*t_k) and a new variance of \mu_k*cos^2(omega_j*t_k).
@@
 Thus, we conclude that for the vast majority of applications, this assumption will hold.
-We did add a statement in the relevant Section 2.1 after Equation 6 to make this more explicit.
+We  added a statement in Section 2.1 after Equation 6 to make this more explicit.
 
@@
 However, in this specific case of dead time, it is ineffective.
-Dead time occurs when after recording of a photon, it cannot detect another photon within a certain time span (it is, effectively, dead).
-This leads to a loss of overall photons, but also to a deviation from the typical Poisson process for photon arrival times to a process that involves a characteristic time scale.
+Dead time occurs when after recording of a photon, the detector cannot detect another photon within a certain time span (it is, effectively, dead).
+This leads to a loss of overall photons, and also to a deviation from the typical Poisson process for photon arrival times to a process that involves a characteristic time scale.
 The resulting periodogram will show an oscillatory pattern over a large range of frequencies (depending on the characteristic time scale of the dead time), and changes both the mean and the variance of the distribution of powers as a function of frequency.
@@
 
-We have added a half-sentence making that explicit.
+We have added a statement making that explicit.
 

### 2017-12-04 paper/referee_response.md
Updated acknowledgments

@@
 In the real word, the phases would be randomly distributed and both positive, negative and near null values would be observed.
+
+and
+
+> 1.
+Re.
+(i) The co-spectrum is real part of product of transforms of the reference and investigated signal.
+At any given frequency if actual oscillation differs in phase by 90/270 degrees from the phase of reference/noise signal component, the result co-spectrum value is null.
+As consequence, there is a range of phases for which even an oscillation with strong S/N remains undetectable.
+A remedy, namely to investigate several phase offsets, wold spoil the claimed for distributions and introduce mutual correlations difficult to analyze.
+This issue is nowhere discussed in the text.
 

### 2017-12-04 paper/cospectra-paper.tex
Updated acknowledgments

@@
 This case is especially relevant for detecting faint X-ray pulsars in detectors heavily affected by instrumental effects, including \nustar, \astrosat\ and \ixpe.
-We show that the statistical distributions of both single and averaged cospectra differ considerably from those for standard power spectra.
+We show that the statistical distributions of both single and averaged cospectra differ considerably from those for standard periodograms.
 While a single cospectrum follows a Laplace distribution exactly, averaged cospectra are approximated by a Gaussian distribution only for more than $\sim\!\!
 30$ averaged segments, dependent on the number of trials.
-We provide an instructive example of a quasi-periodic oscillation in \nustar\ and show that applying standard power spectral statistics leads to underestimated tail probabilities for period detection.
+We provide an instructive example of a quasi-periodic oscillation in \nustar\ and show that applying standard periodogram statistics leads to underestimated tail probabilities for period detection.
 
@@
 Time series analysis is one of the primary ways to understand the physical properties of astronomical object in our universe, from exoplanets and stars to black holes and Active Galactic Nuclei (AGN).
-Fourier analysis, especially through the periodogram\footnote{We distinguish in this paper between the \textit{power spectrum}, which describes the process at the source generating variable time series, and the \textit{periodogram}, which denotes a realization of said power spectrum, i.e.\ the time series we actually observe.}, has long been used to find periodic and quasi-periodic signals as well as characterize the stationary stochastic processes often present in accreting systems.
+Fourier analysis, especially through the periodogram\footnote{We distinguish in this paper between the \textit{power spectrum}, which describes the process at the source generating variable time series, and the \textit{periodogram}, which denotes a realization of said power spectrum, i.e.\ the time series we actually observe.
+\textbf{In line with the signal processing literature, we also use the term \textit{periodogram} or \textit{power spectral density} for the square of the absolute value of Fourier amplitudes derived from an observed light curve}}, has long been used to find periodic and quasi-periodic signals as well as characterize the stationary stochastic processes often present in accreting systems.
 While in principle, the statistics of the periodogram is well understood and characterized in the literature \citep[e.g.][]{vanderklis1989}, the periodogram is often subject to instrumental effects like dead time that change its statistical properties and thus make statistical inference difficult in practice \citep[e.g.][]{Zhang+95}.
@@
 Where in standard analysis, light curves of multiple detectors are summed before Fourier transforming the summed light curve, it is possible to instead Fourier-transform the signal of two independent detectors \textbf{within the same instrument observing the same source strictly simultaneously in the same energy band}.
-The resulting \textbf{co-spectral} powers will be less affected by an instrumental effect called \textit{dead time} (see details in \citealt{Bachetti+15}).
+The resulting \textbf{co-spectral} densities will be less affected by an instrumental effect called \textit{dead time} (see details in \citealt{Bachetti+15}).
 \textbf{In many X-ray detectors, after detection of a photon there is a time interval during which the detector (or in imaging detectors the individual pixel) is blocked from detecting a second photon.
-This time interval is generally called \textit{dead time} (the detector is effectively ``dead'', and incoming photons will not produce a signal) and its characteristic time scale is set by the details of the detector and can range from very short ($2\mu\mathrm{s}$ in the \fermi\ Gamma-Ray Burst Monitor to very long $2.5\mathrm{ms}$ in \nustar).} It leads to frequency-dependent changes in the mean and variance of the statistical distributions governing the periodogram, \textbf{which cannot be mitigated by averaging periodograms of multiple segments}.
-This is relevant for current X-ray missions \textbf{carrying two identical detectors} such as \nustar\ and \astrosat, but will also be relevant to future missions with multiple detectors like \ixpe.
+This time interval is generally called \textit{dead time} (the detector is effectively ``dead'', and incoming photons will not produce a signal) and its characteristic time scale is set by the details of the detector and can range from very short ($2\mu\mathrm{s}$ in the \fermi\ Gamma-Ray Burst Monitor) to very long ($2.5\mathrm{ms}$ in \nustar).} It leads to frequency-dependent changes in the mean and variance of the statistical distributions governing the periodogram, \textbf{which cannot be mitigated by averaging periodograms of multiple segments}.
+This is relevant for current X-ray missions \textbf{carrying at least two identical detectors} such as \nustar\ and \astrosat, but will also be relevant to future missions with multiple detectors like \ixpe.
 This approach has recently been used in \nustar\ studies of millisecond pulsars \citep{ferrigno2017}, Ultraluminous X-ray Sources \citep{bachetti2016} and X-ray binaries \citep{barriere2015,zoghbi2016,ingram2016,huppenkothen2017,stiele2017}.
-\textbf{It is particularly important for instruments like \nustar\ which have a fairly long dead time and thus exhibit dead time effects the power spectrum at frequencies comparable to those where signals are expected, and at fairly low count rates}.
+\textbf{It is particularly important for instruments like \nustar\ which have a fairly long dead time and thus exhibit dead time effects in the periodogram at frequencies comparable to those where signals are expected, and at fairly low count rates}.
 Similarly, the cospectrum of two time series taken with the same instrument, but in different wavelength ranges, can be used to characterize the coherent, \textbf{phase-aligned} variability in both time series as a function of frequency.
 
-While much has been written on the subject of the statistics of cross spectra and time lags \citep[e.g.][]{epitropakis2016}, the derivation of cospectral statistics is notably absent from the astronomy literature, and most publications \textit{utilizing the approach of forming co-spectra from identical, but independent detectors observing the same source} assume that either the $\chi^2_2$ distribution used for standard periodograms or a Gaussian distribution for averaged spectra is appropriate.
+While much has been written on the subject of the statistics of cross spectra and time lags \citep[e.g.][]{epitropakis2016}, the derivation of cospectral statistics is notably absent from the astronomy literature, and most publications \textbf{utilizing the approach of forming co-spectra from identical, but independent detectors observing the same source} assume that either the $\chi^2_2$ distribution used for standard periodograms or a Gaussian distribution for averaged spectra is appropriate.
 That these assumptions are valid has not been shown until now.
@@
 While the white noise is also assumed to be strictly stochastic, the signal to be detected may be either deterministic (in the case of a strictly periodic signal) or stochastic (for a quasi-periodic oscillation).
-We then aim to reject a null hypothesis where the observed cospectral power at a given frequency can be explained by white noise alone.}
+We then aim to reject a null hypothesis where the observed cospectral density at a given frequency can be explained by white noise alone.}
 
@@
 \caption{Distribution of Leahy-normalized cospectral densities (left) and power spectral densities (right), respectively, for the simulated data.
-In dark grey, we show fine-grained histograms of the simulated powers.
-In red we plot the theoretical probability distribution the simulated powers should follow: A Laplace distribution with $\mu=0$ and $\sigma=2$ for the cospectral densities and a $\chi^2$-distribution with $2$ degrees of freedom for the power spectral densities.
-The simulated powers adhere very closely to the theoretical predictions.}
+In dark grey, we show fine-grained histograms of the simulated densities.
+In red we plot the theoretical probability distribution the simulated densities should follow: A Laplace distribution with $\mu=0$ and $\sigma=2$ for the cospectral densities and a $\chi^2$-distribution with $2$ degrees of freedom for the power spectral densities.
+The simulated densities adhere very closely to the theoretical predictions.}
 \label{fig:csdist}
@@
 
-Detection thresholds for cospectra will generally be different from those of classical power spectra, because the Laplace distribution tends to be narrower than the equivalent $\chi^2_2$ distribution for single power spectra.
+Detection thresholds for cospectra will generally be different from those of classical periodograms, because the Laplace distribution tends to be narrower than the equivalent $\chi^2_2$ distribution for single periodograms.
 To show how the distributions and the corresponding detection thresholds differ, we simulated simple Poisson-distributed light curves.
@@
 In order to normalize the cospectrum correctly, we used $2/\sqrt{N_{\mathrm{ph}, x}N_{\mathrm{ph}, y}}$, where $N_{\mathrm{ph}, x}$ and $N_{\mathrm{ph}, y}$ are the number of photons of light curves $\mathbf{x}$ and $\mathbf{y}$, respectively, as prescribed by \citet{Bachetti+15}.
-In this normalization the powers are distributed as $\chi^2_2$ exactly for the periodogram, and following a Laplace distribution with $\mu=0$ and $\sigma = 1$ for the cospectrum.
-In Figure \ref{fig:csdist}, we plot the resulting distribution of powers.
-While the periodogram is only defined for positive values, the Laplace distribution is symmetric around zero, and in general the cospectrum will comprise both positive and negative powers.
+In this normalization the densities are distributed as $\chi^2_2$ exactly for the periodogram, and following a Laplace distribution with $\mu=0$ and $\sigma = 1$ for the cospectrum.
+In Figure \ref{fig:csdist}, we plot the resulting distribution of densities.
+While the periodogram is only defined for positive values, the Laplace distribution is symmetric around zero, and in general the cospectrum will comprise both positive and negative densities.
 It is also immediately visible from Figure \ref{fig:csdist} that the probability of obtaining a certain (positive) noise power is always lower for the Laplace distribution than for the $\chi^2$ distribution.
@@
 In Figure \ref{fig:cs_sim}, we show the cospectral densities along with trial-corrected $0.99$ detection thresholds for both the Laplace and $\chi^2$ distribution.
-If the powers are assumed to follow a $\chi^2$ distribution, as for the periodogram, the candidate at $10 \,\mathrm{Hz}$ would be discounted at the $99\%$ detection threshold, whereas correctly applying the Laplace distribution yields a correct rejection of the null hypothesis at the same detection threshold.
+If the densities are assumed to follow a $\chi^2$ distribution, as for the periodogram, the candidate at $10 \,\mathrm{Hz}$ would be discounted at the $99\%$ detection threshold, whereas correctly applying the Laplace distribution yields a correct rejection of the null hypothesis at the same detection threshold.
 
@@
 The tail probability, or survival function, is defined as $SF(x) = 1 - CDF(x)$.
-The tail probability measures the probability of observing a value $x\geq X$, and is often used for detecting periods in power spectra.
-For the power spectral densities, we plot both the theoretical prediction for the survival function based on the $\chi^2$ distribution (black dashed line), as well as the corrected distribution for power spectra derived in \citet{groth1975} (red solid line).
+The tail probability measures the probability of observing a value $x\geq X$, and is often used for detecting periods in periodograms.
+For the power spectral densities, we plot both the theoretical prediction for the survival function based on the $\chi^2$ distribution (black dashed line), as well as the corrected distribution for periodograms derived in \citet{groth1975} (red solid line).
 For illustrative purposes, we show a single-trial $95\%$ detection threshold for the Laplace distribution (black solid vertical line) and the $\chi^2$ distribution (black dashed vertical line).}
@@
 We also show the 99\% detection threshold, corrected for a number of trials equal to the number of spectral bins, assuming Laplace-distributed data (red solid line) and $\chi^2$-distributed data (red dashed line).
-When the latter distribution is assumed, the periodic signal would not be considered a significant detection, because the $\chi^2$ distribution produces a wider distribution of powers.
+When the latter distribution is assumed, the periodic signal would not be considered a significant detection, because the $\chi^2$ distribution produces a wider distribution of densities.
 Applying the correct Laplace distribution, however, allows for the detection of weaker signals.}
@@
 \includegraphics[width=\textwidth]{avg_dist.png}
-\caption{Histogram of cospectral powers (grey) and the theoretical expectation (red) for three different cases.
-In the right panel, we show the distribution of powers for a single cospectrum, with its expected Laplace distribution from Equation \ref{eqn:laplace}, in the middle and right panel cospectral powers for averaging 10 (middle) and 100 (right) individual cospectra together.
+\caption{Histogram of cospectral densities (grey) and the theoretical expectation (red) for three different cases.
+In the right panel, we show the distribution of densities for a single cospectrum, with its expected Laplace distribution from Equation \ref{eqn:laplace}, in the middle and right panel cospectral densities for averaging 10 (middle) and 100 (right) individual cospectra together.
 In the middle panel, the theoretical expectation of the sampling distribution is given by Equation \ref{eqn:averaged_pdf}, in the right panel by a Gaussian distribution with $\sigma=\sqrt{2/101}$.}
@@
 For practical purposes, evaluating this PDF for averaged spectra above $n\sim 85$ is difficult numerically, because the factorials and exponents in the sum become very large and small, respectively.
-However, as we will show in Section \ref{sec:averaged_detthres} below, \textbf{we expect that for large $n$, the Central Limit Theorem implies that the PDF of averaged cospectral powers tends towards a normal distribution.} We find that in practice, when $n \gtrsim 30$, detection thresholds derived from Equation \ref{eqn:averaged_pdf} provide only a negligible difference over that derived from a normal distribution $N(0, \sqrt{2/(n+1)}$, depending on the significance threshold required and the number of trials.
+However, as we will show in Section \ref{sec:averaged_detthres} below, \textbf{we expect that for large $n$, the Central Limit Theorem implies that the PDF of averaged cospectral densities tends towards a normal distribution.} We find that in practice, when $n \gtrsim 30$, detection thresholds derived from Equation \ref{eqn:averaged_pdf} provide only a negligible difference over that derived from a normal distribution $N(0, \sqrt{2/(n+1)}$, depending on the significance threshold required and the number of trials.
 
@@
 \[
-D = \frac{n(n-1+j)!}{(n-1-j)! (n-1)! 2^{n+j}} \, .
+D = \frac{n(n-1+j)!}{(n-1-j)! j! (n-1)! 2^{n+j}} \, .
 \]
@@
 \includegraphics[width=\textwidth]{avg_pvalues.png}
-\caption{Predictions for single-trial $p$-values as a function of cospectral power, for the simulated data sets (black solid line) compared to the theoretically predicted $p$-values using the CDF derived in Equation \ref{eqn:averaged_cdf} (red dashed line) as well as the survival function of a Gaussian distribution (black dashed line).
+\caption{Predictions for single-trial $p$-values as a function of cospectral density, for the simulated data sets (black solid line) compared to the theoretically predicted $p$-values using the CDF derived in Equation \ref{eqn:averaged_cdf} (red dashed line) as well as the survival function of a Gaussian distribution (black dashed line).
 The exact distribution matches the empirically derived tail probabilities from simulations within the uncertainties of the simulations.
@@
 In order to show the way the probability distribution changes as a function of averaged cospectra, we simulate light curves of $10^{5}$ data points and a mean count rate of $100\,\mathrm{counts}/\mathrm{s}$ consisting of pure white noise.
-We compute $n$ such light curves and average their cospectral powers in order to show the distribution of those powers compared to the expected probability distributions.
-In Figure \ref{fig:avg_dist}, we show the simulated distribution of Leahy-normalized powers, along with the distributions that describe them.
+We compute $n$ such light curves and average their cospectral densities in order to show the distribution of those densities compared to the expected probability distributions.
+In Figure \ref{fig:avg_dist}, we show the simulated distribution of Leahy-normalized power spectral densities, along with the distributions that describe them.
 For a single, non-averaged spectrum, we use the Laplace PDF described in Equation \ref{eqn:laplace}.
-When averaging $10$ cospectra, we use Equation \ref{eqn:averaged_pdf} and show that the theoretical predictions agree with the simulated powers.
-Finally, for an averaged cospectra consisting of 100 individual light curves, Equation \ref{eqn:averaged_pdf} becomes difficult to compute numerically, and we use a Gaussian PDF instead, which describes the distribution of simulated powers well.
+When averaging $10$ cospectra, we use Equation \ref{eqn:averaged_pdf} and show that the theoretical predictions agree with the simulated densities .
+Finally, for an averaged cospectra consisting of 100 individual light curves, Equation \ref{eqn:averaged_pdf} becomes difficult to compute numerically, and we use a Gaussian PDF instead, which describes the distribution of simulated densities well.
 
@@
 The QPO has an rms of $\sim$9\% and a high Q factor of $\sim$40.
-The shape of the periodogram is strongly distorted from the expected flat powerspectrum centred on a constant value of $2$, with deviations of more that $0.5$ in the mean level of the powers common.
+The shape of the periodogram is strongly distorted from the expected flat power spectrum centred on a constant value of $2$, with deviations of more that $0.5$ in the mean level of the densities common.
 In dashed lines, we also show the expected (trial-corrected) $99\%$ confidence level for the periodograms (in the same colour, left panel) as well as the cospectrum (red, right panel).
@@
 These conditions are typically met in most observations taken with X-ray timing instruments like \nustar\ and \rxte\ for problems such as timing of X-ray binaries, AGN and X-ray pulsar searches.
-Note that this assumes, however, that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers.
-Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}).
+Note that this assumes, however, that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers.}
+
+\textbf{Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}).
 Uneven sampling more generally introduces covariances between adjacent frequencies and renders many of the statistical assumptions underpinning period searches using both in the periodogram and the cospectrum invalid.
@@
 \includegraphics[width=\textwidth]{nustar_lc_psd.pdf}
-\caption{\textbf{Upper panel: Light curve of Her X-1 observed with \nustar, observation [XX].
+\caption{\textbf{Upper panel: Light curve of Her X-1 observed with \nustar, observation ID 30002006002.
 The observation totals $32.67\,\mathrm{ks}$, the light curve presented here is produced by combining event data from both FPMA and FPMB detectors.
@@
 Hercules X-1 (Her X-1) is a well-studied persistent X-ray binary pulsars \citep{tananbaum1972} in an X-ray binary comprising the neutron star itself and a stellar companion HZ Herculis \citep{davidsen1972,forman1972,bahcall1972} with a mass of $\sim$$2.2 \,\mathrm{M}_\odot$ \citep{reynolds1997,leahy2014} whose type varies between late-type A and early-type B with orbital phase \citep{anderson1994,cheng1995} .
-The neutron star spins with a period of $P = 1.23\,\mathrm{s}$ [REF?] and the system overall exhibits an orbital period of $P_\mathrm{orb} = 1.7\,\mathrm{days}$ \citep{bahcall1972}, along with super-orbital variations on a $\sim$35-day timescale \citep{giacconi1973,scott1999,igna2011}.
+The neutron star spins with a period of $P = 1.23\,\mathrm{s}$ \citep{tananbaum1972} and the system overall exhibits an orbital period of $P_\mathrm{orb} = 1.7\,\mathrm{days}$ \citep{bahcall1972}, along with super-orbital variations on a $\sim$35-day timescale \citep{giacconi1973,scott1999,igna2011}.
 
@@
 We downloaded the observation directory for observation ID 30002006002 from the HEASARC and used the FTOOL \texttt{barycorr} on the L2 cleaned science event files (file name ending with \verb|01_cl.evt|) to correct the photon arrival times to the solar system barycenter.
-For our analysis, we considered photons from 3 to 79 keV at most 50\arcsec from the nominal position of the source.
+For our analysis, we considered photons from 3 to 79 keV at most 50\arcsec\ from the nominal position of the source, extracted from the two identical \textit{Focal Plane Modules A} and \textit{B} (FPMA and FPMB, respectively) onboard the spacecraft.
 For this work, we used a total of $32.67\,\mathrm{ks}$ of good time intervals (GTIs), only selecting intervals longer than $10\,\mathrm{s}$.
@@
 In order to test our theoretical predictions for the distributions governing (averaged) cospectra, we first produced $3260$ individual cospectra out of each of the $10\,\mathrm{s}$ segments, and for each cospectrum extracted the cospectral densities in the range $50\,\mathrm{Hz}$ to $700\,\mathrm{Hz}$, where we do not expect significant contamination by the pulsar signal and its harmonics, or by the low-frequency red noise component.
-In order to gain sufficient statistics, we plot a histogram of cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ for all $3260$ segments (a total of $21183500$ powers) in Figure \ref{fig:herx1_cs} (lower left panel), along with the theoretically expected Laplace distribution.
+In order to gain sufficient statistics, we plot a histogram of cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ for all $3260$ segments (a total of $21183500$ cospectral densities) in Figure \ref{fig:herx1_cs} (lower left panel), along with the theoretically expected Laplace distribution.
 We find a generally high agreement between the Laplace distribution and the distribution of observed cospectral densities: a Kolmogorov-Smirnov test yields $K= 1.77 \times 10^{-4}$ or a $p$-value of rejecting the null hypothesis that the observed distribution is derived from the theoretically expected Laplace distribution of $p = 0.52$.
@@
 This yields $7$ averaged cospectra.
-As above, we extract cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ and plot the histogram of cospectral densities extracted from all averaged cospectra (910000 powers) in Figure \ref{fig:herx1_cs} (lower right panel).
+As above, we extract cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ and plot the histogram of cospectral densities extracted from all averaged cospectra (910000 cospectral densitites) in Figure \ref{fig:herx1_cs} (lower right panel).
 As with the single cospectrum, the histogram for the averaged cospectral densities matches the theoretical prediction very closely.
@@
 We show that because the Fourier amplitudes being multiplied to derive the cospectrum are now no longer identical (as is the case in the periodogram), the statistical distributions no longer reduce to a simple $\chi^2$ distribution with two degrees of freedom.
-Instead, we find that the powers in a single cospectrum follow a Laplace distribution with a mean of $\mu=0$ and a width of $\sigma=1$.
+Instead, we find that the densities in a single cospectrum follow a Laplace distribution with a mean of $\mu=0$ and a width of $\sigma=1$.
 This has important consequences for period detection.
@@
 As an example, we have simulated how a QPO would appear in \nustar\ in the presence of dead time, and have shown that the shape of the periodogram is strongly distorted, whereas that of the cospectrum is not (for a longer introduction into the cospectrum and how it can be used in the presence of dead time, see also \citealt{Bachetti+15}).
-The significance of the QPO is difficult to assess in the periodogram, because of the non-linearities introduced into the power spectrum by the variable dead time.
-The standard $\chi^2_2$ distributions may either overestimate or underestimate the significance, depending on the shape of the power spectrum at a given frequency, adding complexity to the detection process in the form of finding a non-linear model for the dead time.
+The significance of the QPO is difficult to assess in the periodogram, because of the non-linearities introduced by the variable dead time.
+The standard $\chi^2_2$ distributions may either overestimate or underestimate the significance, depending on the shape of the underlying power spectrum and its modification due to dead time at a given frequency, adding complexity to the detection process in the form of finding a non-linear model for the dead time.
 The cospectrum, on the other hand, only requires an estimate of the local variance in order to use the equations derived above, making it a far more convenient choice for periodicity detection.
@@
 \paragraph{Acknowledgements}
-The authors thank Thomas Laetsch for helpful suggestions regarding the mathematical derivations.
+The authors thank the anonymous referee for their helpful comments.
+The authors also thank Thomas Laetsch for helpful suggestions regarding the mathematical derivations, and Peter Bult for spotting a mathematical error.
 DH is supported by the James Arthur Postdoctoral Fellowship and the Moore-Sloan Data Science Environment at New York University. 

### 2017-12-05 paper/cospectra-paper.tex
Added KS test for averaged cospectra

@@
 In red, we show the $p = 10^{-3}$ detection threshold, corrected for $6999$ trials (equivalent to the number of frequencies in the periodogram).
-Lower left:  histogram (grey) of the cospectral densities collected from 3260 individual cospectra between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$.
+Lower left:  histogram (grey) of the cospectral densities collected from 3260 individual cospectra between $50\,\mathrm{Hz}$ and $400\,\mathrm{Hz}$.
 In red, we also show the theoretically predicted Laplace distribution.
 Lower right: in black, the histogram of cospectral densities collected from $7$ averaged cospectra, each created from $15$ light curve segments of $200\,\mathrm{s}$ duration.
-Again, only cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ were used in this figure.
+Again, only cospectral densities between $50\,\mathrm{Hz}$ and $400\,\mathrm{Hz}$ were used in this figure.
 In red, we show the theoretical probability density function defined in Equation \ref{eqn:averaged_pdf} with $n = 15$.
@@
 The neutron star spins with a period of $P = 1.23\,\mathrm{s}$ \citep{tananbaum1972} and the system overall exhibits an orbital period of $P_\mathrm{orb} = 1.7\,\mathrm{days}$ \citep{bahcall1972}, along with super-orbital variations on a $\sim$35-day timescale \citep{giacconi1973,scott1999,igna2011}.
-
+q
 Her X-1 was observed by the Nuclear Spectroscopic Telescope Array (\nustar) multiple times in the last years. 
@@
 
-In order to test our theoretical predictions for the distributions governing (averaged) cospectra, we first produced $3260$ individual cospectra out of each of the $10\,\mathrm{s}$ segments, and for each cospectrum extracted the cospectral densities in the range $50\,\mathrm{Hz}$ to $700\,\mathrm{Hz}$, where we do not expect significant contamination by the pulsar signal and its harmonics, or by the low-frequency red noise component.
-In order to gain sufficient statistics, we plot a histogram of cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ for all $3260$ segments (a total of $21183500$ cospectral densities) in Figure \ref{fig:herx1_cs} (lower left panel), along with the theoretically expected Laplace distribution.
-We find a generally high agreement between the Laplace distribution and the distribution of observed cospectral densities: a Kolmogorov-Smirnov test yields $K= 1.77 \times 10^{-4}$ or a $p$-value of rejecting the null hypothesis that the observed distribution is derived from the theoretically expected Laplace distribution of $p = 0.52$.
+In order to test our theoretical predictions for the distributions governing (averaged) cospectra, we first produced $3260$ individual cospectra out of each of the $10\,\mathrm{s}$ segments, and for each cospectrum extracted the cospectral densities in the range $50\,\mathrm{Hz}$ to $400\,\mathrm{Hz}$, where we do not expect significant contamination by the pulsar signal and its harmonics, or by the low-frequency red noise component.
+In order to gain sufficient statistics, we plot a histogram of cospectral densities between $50\,\mathrm{Hz}$ and $400\,\mathrm{Hz}$ for all $3260$ segments (a total of $11406500$ cospectral densities) in Figure \ref{fig:herx1_cs} (lower left panel), along with the theoretically expected Laplace distribution.
+We find a generally high agreement between the Laplace distribution and the distribution of observed cospectral densities: a two-sided Kolmogorov-Smirnov test yields $K= 2.33 \times 10^{-4}$ or a $p$-value of rejecting the null hypothesis that the observed distribution is derived from the theoretically expected Laplace distribution of $p = 0.57$.
 
@@
 This yields $7$ averaged cospectra.
-As above, we extract cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ and plot the histogram of cospectral densities extracted from all averaged cospectra (910000 cospectral densitites) in Figure \ref{fig:herx1_cs} (lower right panel).
+As above, we extract cospectral densities between $50\,\mathrm{Hz}$ and $400\,\mathrm{Hz}$ and plot the histogram of cospectral densities extracted from all averaged cospectra (910000 cospectral densitites) in Figure \ref{fig:herx1_cs} (lower right panel).
 As with the single cospectrum, the histogram for the averaged cospectral densities matches the theoretical prediction very closely.
+We compute the two-sided Kolmogorov-Smirnov statistic for the histogram of averaged cospectral densities, compared to the theoretical expected distribution, as$k = 0.01$, equivalent to a p-value of $p = 0.19$.
+This indicates that there is no evidence in the data that would argue against the observed cospectral densities being drawn from the distribution in Equation \ref{eqn:averaged_cdf}.
 \end{bf}

### 2017-12-10 paper/cospectra-paper.tex
Some minor additions

@@
 
-In the following, we first simulate a QPO as expected to be observed in a typical \nustar\ observation, and then show similarly real \nustar\ data containing a signal.
+In the following, we first simulate a QPO as expected to be observed in a typical \nustar\ observation, and then show similarly real \nustar\ data containing a coherent pulsed signal.
 
@@
 To this constant background we add a quasi-periodic oscillation with a period of $5\,\mathrm{ms}$, a fractional rms amplitude of $f_\mathrm{rms} = 0.15$ and phases randomized using a normal distribution with a mean of $0$ and a width of $\sigma_\mathrm{qpo} = 0.01$.
-We then simulated photon events using rejection sampling from this light curve using the software package \textit{stingray}\footnote{\url{https://github.com/StingraySoftware/stingray}}, running the function \texttt{stingray.Eventlist.simulate\_times()} twice in order to produce two light curves that are statistically independent, but have the same signal and properties, as we would expect from an instrument with two independent detectors observing the same object.
+After producing this light curve, the procedure is similar to that followed by \citet{Bachetti+15}.
+We simulated photon events using rejection sampling from this light curve using the software package \textit{stingray}\footnote{\url{https://github.com/StingraySoftware/stingray}}, running the function \texttt{stingray.Eventlist.simulate\_times()} twice in order to produce two light curves that are statistically independent, but have the same signal and properties, as we would expect from an instrument with two independent detectors observing the same object.
 Subsequently we simulated variable deadtime for both light curves with an average time scale of $2.5 \,\mathrm{ms}$ as commonly seen in \nustar\ data \citep{Bachetti+15} using \textit{HENDRICS}\footnote{\url{https://github.com/StingraySoftware/HENDRICS}} \citep{bachetti2015b}.

### 2017-12-18 paper/cospectra-paper.tex
New section on white noise simulations with dead time

@@
 Fourier analysis, especially through the periodogram\footnote{We distinguish in this paper between the \textit{power spectrum}, which describes the process at the source generating variable time series, and the \textit{periodogram}, which denotes a realization of said power spectrum, i.e.\ the time series we actually observe.
-\textbf{In line with the signal processing literature, we also use the term \textit{periodogram} or \textit{power spectral density} for the square of the absolute value of Fourier amplitudes derived from an observed light curve}}, has long been used to find periodic and quasi-periodic signals as well as characterize the stationary stochastic processes often present in accreting systems.
+\textbf{In line with the signal processing literature, we also use the term \textit{periodogram} or \textit{power spectral density} for the square of the absolute value of Fourier amplitudes derived from an observed light curve.
+We use the terms \textit{cospectrum} or \textit{cospectral density} to denote the real component of the cross-spectrum, i.e.\ the result of a multiplication of the Fourier amplitudes of one light curve with the complex conjugate of the Fourier amplitudes of a second light curve (see also Section \ref{sec:whitenoise_cospectra} for an exact definition)}}, has long been used to find periodic and quasi-periodic signals as well as characterize the stationary stochastic processes often present in accreting systems.
 While in principle, the statistics of the periodogram is well understood and characterized in the literature \citep[e.g.][]{vanderklis1989}, the periodogram is often subject to instrumental effects like dead time that change its statistical properties and thus make statistical inference difficult in practice \citep[e.g.][]{Zhang+95}.
@@
 
-Within astronomy, it has recently been recognized as one approach to mitigate instrumental effects by making use of the independent detectors present in some telescopes.
-Where in standard analysis, light curves of multiple detectors are summed before Fourier transforming the summed light curve, it is possible to instead Fourier-transform the signal of two independent detectors \textbf{within the same instrument observing the same source strictly simultaneously in the same energy band}.
-The resulting \textbf{co-spectral} densities will be less affected by an instrumental effect called \textit{dead time} (see details in \citealt{Bachetti+15}).
+Within astronomy, it has recently been recognized as one approach to mitigate instrumental effects by making use of the independent detectors present in some telescopes, in particular an effect called \textit{dead time}.
 \textbf{In many X-ray detectors, after detection of a photon there is a time interval during which the detector (or in imaging detectors the individual pixel) is blocked from detecting a second photon.
 This time interval is generally called \textit{dead time} (the detector is effectively ``dead'', and incoming photons will not produce a signal) and its characteristic time scale is set by the details of the detector and can range from very short ($2\mu\mathrm{s}$ in the \fermi\ Gamma-Ray Burst Monitor) to very long ($2.5\mathrm{ms}$ in \nustar).} It leads to frequency-dependent changes in the mean and variance of the statistical distributions governing the periodogram, \textbf{which cannot be mitigated by averaging periodograms of multiple segments}.
-This is relevant for current X-ray missions \textbf{carrying at least two identical detectors} such as \nustar\ and \astrosat, but will also be relevant to future missions with multiple detectors like \ixpe.
-This approach has recently been used in \nustar\ studies of millisecond pulsars \citep{ferrigno2017}, Ultraluminous X-ray Sources \citep{bachetti2016} and X-ray binaries \citep{barriere2015,zoghbi2016,ingram2016,huppenkothen2017,stiele2017}.
+
+Where in standard analysis, light curves of multiple detectors are summed before Fourier transforming the summed light curve, it is possible to instead Fourier-transform the signal of two independent detectors \textbf{within the same instrument observing the same source strictly simultaneously in the same energy band}.
+The resulting \textbf{co-spectral} densities will be less affected by dead time (see details in \citealt{Bachetti+15}), because the latter is introduced in each detector independently and tends to cancel out in the cospectrum.
+This approach relevant for current X-ray missions \textbf{carrying at least two identical detectors} such as \nustar\ and \astrosat, but will also be relevant to future missions with multiple detectors like \ixpe.
+It has recently been used in \nustar\ studies of millisecond pulsars \citep{ferrigno2017}, Ultraluminous X-ray Sources \citep{bachetti2016} and X-ray binaries \citep{barriere2015,zoghbi2016,ingram2016,huppenkothen2017,stiele2017}.
 \textbf{It is particularly important for instruments like \nustar\ which have a fairly long dead time and thus exhibit dead time effects in the periodogram at frequencies comparable to those where signals are expected, and at fairly low count rates}.
@@
 \textbf{We note that the results below hold only for cases where the signal to be detected is phase-aligned in both light curves.
-This must be true by definition for analyses utilizing data from independent, identical detectors observing the same source simultaneously, but might not strictly be true for observations of the same source e.g.\ in different energy bands.
+This must be true by definition for analyses utilizing data from independent, identical detectors observing the same source simultaneously (unless detectors are so far apart that light travel time effects might be an issue and the data is not barycentred to the gravitational centre of the solar system), but might not strictly be true for observations of the same source e.g.\ in different energy bands.
 We caution the reader that for signals that are not phase-aligned, the resulting power may decrease significantly and result in a non-detectable signal.}
@@
 
-\textbf{In the following, we will consider the example of detecting a strictly periodic or very narrow quasi-periodic signal in the presence of simple white noise, as is commonly the case in pulsar searches in X-ray data.
-We assume that both the white noise and the periodic signal are wide-sense stationary and the light curve is evenly sampled.
+\textbf{In the following, we will consider the example of detecting a strictly periodic or very narrow quasi-periodic signal in the presence of simple white noise, as is commonly the case for example in pulsar searches in X-ray data.
+We assume that both the white noise and the periodic signal are wide-sense stationary (i.e.
+the mean and the autocovariance of the time series do not change with time) and the light curve is evenly sampled simultaneously in two identical detectors.
 While the white noise is also assumed to be strictly stochastic, the signal to be detected may be either deterministic (in the case of a strictly periodic signal) or stochastic (for a quasi-periodic oscillation).
@@
 \noindent For \textbf{evenly sampled}, normally distributed light curves (and indeed for uncertainties coming from a wide range of statistical distributions, including the Poisson distribution, if $N$ is large), the real and imaginary amplitude components are distributed as $A_{xj}, B_{xj} \sim \mathcal{N}(0, \sigma_x^2)$ with $\sigma_x =  \sqrt{\frac{\sum_{k=1}^{N}{x_k}}{2}}$ and  $A_{yj}, B_{yj} \sim \mathcal{N}(0, \sigma_y^2)$ with $\sigma_y = \sqrt{\frac{\sum_{k=1}^{N}{y_k}}{2}}$.
-\textbf{Note that these distributions for the Fourier amplitudes only hold if the underlying process producing the observations is stochastic and wide-sense stationary (like e.g.\ the white noise case considered here) as is commonly observed in a wide range of astrophysical sources including black hole X-ray binaries and AGN.
-For strictly deterministic processes, the distributions of the Fourier amplitudes will not be centered on $\mu = 0$, and the distributions below will not be correct.}
+\textbf{Note that these distributions for the Fourier amplitudes only hold if the underlying process producing the observations is stochastic and wide-sense stationary.
+This includes many processes commonly observed in astrophysical sources such as white noise observed from a constant background, as well as red noise processes such as shot noise and other (broken) power-law power spectra often seen in AGN, and quasi-periodic oscillations with stochastic variations in amplitude and period of the observed signal, commonly observed in black hole X-ray binaries.
+For strictly deterministic processes (e.g.\ strictly periodic variations), the distributions of the Fourier amplitudes will not be centered on $\mu = 0$, and the distributions below will not be correct.}
 
@@
 \end{equation}
-but since $\mu_x = \mu_y = 0$ for Fourier amplitudes, this reduces to
+but since $\mu_x = \mu_y = 0$ for the Fourier amplitudes \textbf{of stationary stochastic processes}, this reduces to
 \begin{equation}
@@
 The light curves have an identical mean count rate of $10^{6} \, \mathrm{counts/s}$, corresponding to $10$ counts per bin.
-In order to simulate typical measurement uncertainties, we sampled from a Poisson distribution for each time bin with a rate parameter $\lambda = 10$, corresponding to the average counts per bin.
+In order to simulate typical measurement uncertainties in X-ray detectors, we sampled from a Poisson distribution for each time bin with a rate parameter $\lambda = 10$, corresponding to the average counts per bin.
 
@@
 
+\section{\textbf{Caveats}}
+
+\textbf{All caveats applying more generally to Fourier analysis and to periodograms specifically also apply to cospectra.
+In particular, as for the $\chi^2$ distribution used for period searches in periodograms, the distributions derived here assume that the underlying process is wide-sense stationary and the light curves are simultaneously and evenly sampled.
+These conditions are typically met in most observations taken with X-ray timing instruments like \nustar\ and \rxte\ for problems such as timing of X-ray binaries, AGN and X-ray pulsar searches, but this may not be the case for e.g.\ optical observations of main-sequence or binary stars.
+Note that we also implicitly assume that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers, both of which may not be stationary.}
+
+\textbf{Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}).
+Uneven sampling, on the other hand, introduces covariances between adjacent frequencies and renders many of the statistical assumptions underpinning period searches using both in the periodogram and the cospectrum invalid.
+For unevenly sampled data, other methods that do not depend on a regular sampling pattern must be employed (for example the Lomb-Scargle periodogram \citep{lomb1976, scargle1982} or the Bayesian period search in \citet{gregory1992}) and dead-time effects must be forward-simulated.}
+
+\textbf{An additional assumption is the phase-alignment of the periodic signal to be detected in both light curves.
+This is always true when the same source is observed simultaneously in the same energy range in independent, but identical detectors on the same spacecraft, as is common practice for instruments like \nustar\, \astrosat\ and \rxte.
+Because the phase of the periodic signal can more generally be energy-dependent, we caution the reader to be careful when producing cospectra of light curves taken for example in different energy ranges.
+Similarly, the phase of a quasi-periodic signal can shift with times, thus observations at different points in time, or of different sources must be analyzed similarly carefully.
+In all these cases, period searches will only be sensitive to signals that are phase-aligned, and will in the worst case be undetectable if the phase-shift is 90 or 270 degrees.
+}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{whitenoise_deadtime.png}
+\caption{\textbf{Left panel: Fourier products derived from white noise light curves heavily affected by dead time.
+There is a strong, frequency-dependent modulation in the periodogram (orange) introduced by dead time effects, rendering standard tests for periodicities unreliable.
+The cospectrum (dark grey) is less affected overall, but there remains a frequency-dependent variation in the local variance, with a narrower distribution governing the cospectral densities at low frequencies, and a wider distribution e.g.\ between 200 and 400 Hz.
+Applying the FAD correction (blue) produces a cospectrum that follows the distributions defined in this paper very closely, allowing for reliable periodicity testing.
+Right panel: histograms of cospectral densities derived from $1000$ simulated, dead time-affected cospectra (dark grey) and their FAD-corrected equivalents (blue).
+The dead time-affected cospectra overall have a narrower distribution (but note that this is integrated over all frequencies, thus the frequency dependence is washed out), whereas the corrected cospectra follow the expected distribution (red).
+}}
+\label{fig:wn_dt}
+\end{center}
+\end{figure*}
+
 \begin{figure*}
@@
 
-\section{\textbf{Caveats}}
+\section{Example: \nustar}
+\label{sec:nustarqpo}
 
-\textbf{All caveats applying more generally to Fourier analysis and to periodograms specifically also apply to cospectra.
-In particular, as for the $\chi^2$ distribution used for period searches in periodograms, the distributions derived here assume that the underlying process is wide-sense stationary and the light curves are simultaneously and evenly sampled.
-These conditions are typically met in most observations taken with X-ray timing instruments like \nustar\ and \rxte\ for problems such as timing of X-ray binaries, AGN and X-ray pulsar searches.
-Note that this assumes, however, that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers.}
+\textbf{In the following, we will consider three more realistic examples in turn.
+First, we will consider simulations of white noise light curves with strong dead time effects.
+We will then provide more realistic simulations of a QPO as expected to be observed in a typical \nustar\ observation, and finally present an example of real \nustar\ data containing a coherent pulsed signal from the X-ray neutron star source Hercules X-1.}
 
-\textbf{Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}).
-Uneven sampling more generally introduces covariances between adjacent frequencies and renders many of the statistical assumptions underpinning period searches using both in the periodogram and the cospectrum invalid.
-For unevenly sampled data, other methods that do not depend on a regular sampling pattern must be employed (for example the Lomb-Scargle periodogram \citep{lomb1976, scargle1982} or the Bayesian period search in \citet{gregory1992}).}
+\subsection{Simulated Light Curves With Dead Time}
 
-\textbf{An additional assumption is the phase-alignment of the periodic signal to be detected in both light curves.
-This is always true when the same source is observed simultaneously in the same energy range in independent, but identical detectors on the same spacecraft, as is common practice for instruments like \nustar\, \astrosat\ and \rxte.
-Because the phase of the periodic signal can more generally be energy-dependent, we caution the reader to be careful when producing cospectra of light curves taken for example in different energy ranges.
-Similarly, the phase of a quasi-periodic signal can shift with times, thus observations at different points in time, or of different sources must be analyzed similarly carefully.
-In all these cases, period searches will only be sensitive to signals that are phase-aligned, and will in the worst case be undetectable if the phase-shift is 90 or 270 degrees.
-}
+\begin{bf}
+Here, we simulate simple white noise light curves and apply a typical \nustar-like dead time of $2.5\mathrm{ms}$ in order to explore how the distributions defined above work under more realistic conditions.
+We use the relatively extreme case of the Crab pulsar, one of the brightest variable X-ray sources in the sky [REF].
+We simulate event files for an observation of $T = 100\,\mathrm{s}$ and a constant count rate of $c = 400 \mathrm{counts}/\mathrm{s}$.
+Incident photon events are generated by picking $N=cT = 4 \times 10^{4}$ times of arrivals randomly from a uniform distribution between the start and end time, independently for each of the two \nustar\ detectors.
+We then apply dead time to both sets of photon events by excluding all photons that arrive within $2.5\mathrm{ms}$ of the previous photon arrival in each detector.
+For a count rate of $400\mathrm{counts}/\mathrm{s}$ and a \nustar-like dead time, this corresponds to a photon loss of $\sim$50\%.
+The resulting light curves are binned at a time resolution of $9.766 \times 10^{-4}\,\mathrm{s}$ (corresponding to a Nyquist frequency of $\nu_{\mathrm{Nyquist}} = 512 \,\mathrm{Hz}$) and averaged periodograms of each individual light curve, of the summed light curves, and the averaged cospectrum are created using segments of $t_\mathrm{seg} = 10\,\mathrm{s}$.
+We repeated this process 1000 times in order to gain sufficient simulated cospectral densities for statistical comparisons.
+
+In Figure \ref{fig:wn_dt} (left panel), we show the periodogram, cospectrum and the corresponding probability distribution of cospectral densities of these simulations.
+In the periodogram, we see a large frequency-dependent deviation from the expected white noise level of $2$, which is difficult to correct for in practice and introduces a large frequency-dependent bias into any standard tests for periodicities that utilize the periodogram.
+The cospectrum, on the other hand, does not suffer from distortions of the underlying power spectrum, but does show frequency-dependent variations in the local variance (see also \citet{Bachetti+15} for a more in-depth explanation of this effect).
+This indicates that the distributions defined in this paper cannot be applied to dead-time affected data without correction.
+Here, we follow the suggestion of \citet{bachetti2017} and use so-called \textit{Fourier Amplitude Differences} (FAD) in order to correct the cospectrum.
+In short, this technique rests on the idea that while variability intrinsic to the source will produce Fourier amplitudes with the same real and imaginary components, both dead time and Photon counting noise are independent between detectors, and will thus be uncorrelated between Fourier amplitudes derived from two identical, independent detectors.
+The \textit{difference} between Fourier amplitudes of one detector and another will hence effectively remove source-intrinsic variability and only contain dead-time related effects.
+One can then model the latter in the Fourier amplitude differences, and use this model to correct the original cospectrum (see \citet{bachetti2017} for details).
+
+We corrected the cospectra using the FAD method and subsequently compared the empirical distributions derived from the simulated cospectra to the theoretical distribution defined in Section \ref{sec:averaged_detthres} (Figure \ref{fig:wn_dt}, right panel).
+We find that the FAD-corrected cospectra follow the derived distributions very closely even in this case involving an exceptionally high degree of dead time, and conclude that for most realistic scenarios, the distributions arrived here can be used to construct unbiased estimators for periodicities, assuming that the FAD correction is applied to the cospectra first.
+
+\end{bf}
 
@@
 
-\section{Example: A (Q)PO in \nustar}
-\label{sec:nustarqpo}
+\subsection{A Simulated QPO in \nustar}
 
-In the following, we first simulate a QPO as expected to be observed in a typical \nustar\ observation, and then show similarly real \nustar\ data containing a coherent pulsed signal.
-
-\subsection{Simulated Data}
 In order to show the difference of the detection limits with the cospectrum and the power spectrum, we show how a QPO at 200 Hz from a very bright source would appear in \nustar. 
@@
 Note that in all three cases we produced averaged periodograms and co-spectra by splitting the light curves into 600 segments of $5\,\mathrm{s}$ length each in order to suppress the variance in the powers and show the effects of dead time more clearly.
+\textbf{The cospectra were FAD-corrected after averaging.}
 
@@
 
-\subsection{\nustar\ Observations}
+\subsection{\nustar\ Observations of Her X-1}
 
@@
 The neutron star spins with a period of $P = 1.23\,\mathrm{s}$ \citep{tananbaum1972} and the system overall exhibits an orbital period of $P_\mathrm{orb} = 1.7\,\mathrm{days}$ \citep{bahcall1972}, along with super-orbital variations on a $\sim$35-day timescale \citep{giacconi1973,scott1999,igna2011}.
-q
+
 Her X-1 was observed by the Nuclear Spectroscopic Telescope Array (\nustar) multiple times in the last years. 

### 2017-12-18 paper/cospectra-paper.tex
Added some more text

@@
 
-As an example, we have simulated how a QPO would appear in \nustar\ in the presence of dead time, and have shown that the shape of the periodogram is strongly distorted, whereas that of the cospectrum is not (for a longer introduction into the cospectrum and how it can be used in the presence of dead time, see also \citealt{Bachetti+15}).
+As an example, we have simulated how both a bright constant light curve and a QPO would appear in \nustar\ in the presence of dead time, and have shown that the shape of the periodogram is strongly distorted, whereas that of the cospectrum is not (for a longer introduction into the cospectrum and how it can be used in the presence of dead time, see also \citealt{Bachetti+15}).
 The significance of the QPO is difficult to assess in the periodogram, because of the non-linearities introduced by the variable dead time.
@@
 The cospectrum, on the other hand, only requires an estimate of the local variance in order to use the equations derived above, making it a far more convenient choice for periodicity detection.
+\textbf{Using a straightforward correction using the differences between Fourier amplitudes derived from two independent detectors, cospectra can be corrected such that they follow the distributions derived here very closely.}
 
@@
 For sources observed with \nustar\ and other instruments with a similar set-up of at least two redundant, identical detectors, one can take advantage of the independent, simultaneous observations in two detectors to form a cospectrum.
-While any (quasi-)periodic signal will appear weaker in the cospectrum than in the periodogram (even in the case, as presented here, where the signals in the two detectors are exactly in phase), it has the substantial advantage that it is unaffected by dead time and other similar detector effects.
-We show that the cospectrum of the observations observed with \nustar\ follow the expected theoretical distributions very closely.}
+While any (quasi-)periodic signal will appear weaker in the cospectrum than in the periodogram (even in the case, as presented here, where the signals in the two detectors are exactly in phase), it has the substantial advantage that it is less affected by dead time and other similar detector effects.
+We show that the cospectrum of the observations observed with \nustar\ follow the expected theoretical distributions very closely once the FAD correction is applied.}
 

### 2017-12-19 paper/cospectra-paper.tex
Correct reference to equation 7
Updated text after discussion on telecon

@@
 
-Within astronomy, it has recently been recognized as one approach to mitigate instrumental effects by making use of the independent detectors present in some telescopes, in particular an effect called \textit{dead time}.
+Within astronomy, and in particular the field of X-ray timing, it has recently been recognized \textbf{ as a powerful solution to two different problems.
+One problem is the reliable estimation of the variability contained in certain features of the power spectrum like quasi-periodic oscillations (QPOs) via the fractional rms amplitude.
+The rms amplitude crucially depends on reliable estimation of the Poisson noise level, which is sometimes difficult to calculate in practice.
+For instruments with two independent, identical detectors, the cospectrum of the light curves measured in each detector, respectively, will not contain any white noise contribution, since Poisson counting statistics are effects local to the instrument, and thus the observed data sets are independently sampled in each detector.
+This makes estimation of source-intrinsic variability much more reliable than comparable measurements using the periodogram.}
+
+The cospectrum is also effectively used as as one approach to mitigate instrumental effects , in particular an effect called \textit{dead time}.
 \textbf{In many X-ray detectors, after detection of a photon there is a time interval during which the detector (or in imaging detectors the individual pixel) is blocked from detecting a second photon.
@@
 
+\textbf{Note that for light curves affected by dead time, the resulting cospectrum will still follow the Laplace distribution above, but with a variable variance that changes as a function of frequency \citep{Bachetti+15}.
+In practice, this issue can be straightforwardly corrected using the differences in Fourier amplitudes derived from the light curves of two detectors, and \citet{bachetti2017} show that the corrected cospectrum will closely follow the Laplace distribution derived here, allowing for unbiased significance tests for periodicity detection.}
+
 \begin{figure}
@@
 
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{whitenoise_deadtime.png}
-\caption{\textbf{Left panel: Fourier products derived from white noise light curves heavily affected by dead time.
+\textbf{In realistic applications, in particularly where dead time is a major concern, the cospectrum will show these detector effects in the form of a frequency-dependent variation in the local variance.
+It is imperative that this should be corrected before using the distributions used here by applying the Fourier Amplitude Differencing (FAD) technique \citep{bachetti2017}.
+This method, while powerful, is not without caveats.
+In particularly, FAD-corrected cospectra tend to overestimate the integrated rms when both count rate and variability amplitude are both very high.
+This should be accounted for when deriving estimates of the fractional rms amplitude.}
+
+%\begin{figure*}
+%\begin{center}
+%\includegraphics[width=\textwidth]{whitenoise_deadtime.png}
+%\caption{\textbf{Left panel: Fourier products derived from white noise light curves heavily affected by dead time.
 There is a strong, frequency-dependent modulation in the periodogram (orange) introduced by dead time effects, rendering standard tests for periodicities unreliable.
@@
 The dead time-affected cospectra overall have a narrower distribution (but note that this is integrated over all frequencies, thus the frequency dependence is washed out), whereas the corrected cospectra follow the expected distribution (red).
-}}
-\label{fig:wn_dt}
-\end{center}
-\end{figure*}
+%}}
+%\label{fig:wn_dt}
+%\end{center}
+%\end{figure*}
 
@@
 
-\textbf{In the following, we will consider three more realistic examples in turn.
-First, we will consider simulations of white noise light curves with strong dead time effects.
-We will then provide more realistic simulations of a QPO as expected to be observed in a typical \nustar\ observation, and finally present an example of real \nustar\ data containing a coherent pulsed signal from the X-ray neutron star source Hercules X-1.}
+\textbf{In the following, we will consider two more realistic examples in turn.
+First, we will consider a more realistic simulation of a QPO as expected to be observed in a typical \nustar\ observation.
+Subsequently, we present an example of real \nustar\ data containing a coherent pulsed signal from the X-ray neutron star source Hercules X-1.}
 
-\subsection{Simulated Light Curves With Dead Time}
+%\subsection{Simulated Light Curves With Dead Time}
 
-\begin{bf}
-Here, we simulate simple white noise light curves and apply a typical \nustar-like dead time of $2.5\mathrm{ms}$ in order to explore how the distributions defined above work under more realistic conditions.
+%\begin{bf}
+%Here, we simulate simple white noise light curves and apply a typical \nustar-like dead time of $2.5\mathrm{ms}$ in order to explore how the distributions defined above work under more realistic conditions.
 We use the relatively extreme case of the Crab pulsar, one of the brightest variable X-ray sources in the sky [REF].
@@
 
-In Figure \ref{fig:wn_dt} (left panel), we show the periodogram, cospectrum and the corresponding probability distribution of cospectral densities of these simulations.
+%In Figure \ref{fig:wn_dt} (left panel), we show the periodogram, cospectrum and the corresponding probability distribution of cospectral densities of these simulations.
 In the periodogram, we see a large frequency-dependent deviation from the expected white noise level of $2$, which is difficult to correct for in practice and introduces a large frequency-dependent bias into any standard tests for periodicities that utilize the periodogram.
 The cospectrum, on the other hand, does not suffer from distortions of the underlying power spectrum, but does show frequency-dependent variations in the local variance (see also \citet{Bachetti+15} for a more in-depth explanation of this effect).
-This indicates that the distributions defined in this paper cannot be applied to dead-time affected data without correction.
+%This indicates that the distributions defined in this paper cannot be applied to dead-time affected data without correction.
 Here, we follow the suggestion of \citet{bachetti2017} and use so-called \textit{Fourier Amplitude Differences} (FAD) in order to correct the cospectrum.
@@
 
-We corrected the cospectra using the FAD method and subsequently compared the empirical distributions derived from the simulated cospectra to the theoretical distribution defined in Section \ref{sec:averaged_detthres} (Figure \ref{fig:wn_dt}, right panel).
+%We corrected the cospectra using the FAD method and subsequently compared the empirical distributions derived from the simulated cospectra to the theoretical distribution defined in Section \ref{sec:averaged_detthres} (Figure \ref{fig:wn_dt}, right panel).
 We find that the FAD-corrected cospectra follow the derived distributions very closely even in this case involving an exceptionally high degree of dead time, and conclude that for most realistic scenarios, the distributions arrived here can be used to construct unbiased estimators for periodicities, assuming that the FAD correction is applied to the cospectra first.
 
-\end{bf}
+%\end{bf}
 
@@
 Note that in all three cases we produced averaged periodograms and co-spectra by splitting the light curves into 600 segments of $5\,\mathrm{s}$ length each in order to suppress the variance in the powers and show the effects of dead time more clearly.
-\textbf{The cospectra were FAD-corrected after averaging.}
+\textbf{As mentioned above, cospectra with dead time are subject to frequency-dependent changes in variance, and we thus corrected the simulated cospectra using the FAD technique.}
 
@@
 Because the periodic signal generated by the pulsar dominates the cospectrum, we plot the cospectrum of a short segment of $10\,\mathrm{s}$ duration starting at $\mathrm{MET} = 85740109.54169$ for illustrative purpose to highlight the noise properties.
-Unlike the periodogram, the cospectrum is not affected by dead time, and the statistical distributions defined in Section \ref{sec:whitenoise_cospectra} apply (see also Figure \ref{fig:herx1_cs}, lower left panel and below).
+Unlike the periodogram, the cospectrum is not modulated by dead time, and the statistical distributions defined in Section \ref{sec:whitenoise_cospectra} apply once the cospectra are corrected using the FAD technique (see also Figure \ref{fig:herx1_cs}, lower left panel and below).
 We also show the trial-corrected detection threshold under the null hypothesis that the cospectrum consists solely of white noise.
@@
 
-In order to test our theoretical predictions for the distributions governing (averaged) cospectra, we first produced $3260$ individual cospectra out of each of the $10\,\mathrm{s}$ segments, and for each cospectrum extracted the cospectral densities in the range $50\,\mathrm{Hz}$ to $400\,\mathrm{Hz}$, where we do not expect significant contamination by the pulsar signal and its harmonics, or by the low-frequency red noise component.
-In order to gain sufficient statistics, we plot a histogram of cospectral densities between $50\,\mathrm{Hz}$ and $400\,\mathrm{Hz}$ for all $3260$ segments (a total of $11406500$ cospectral densities) in Figure \ref{fig:herx1_cs} (lower left panel), along with the theoretically expected Laplace distribution.
-We find a generally high agreement between the Laplace distribution and the distribution of observed cospectral densities: a two-sided Kolmogorov-Smirnov test yields $K= 2.33 \times 10^{-4}$ or a $p$-value of rejecting the null hypothesis that the observed distribution is derived from the theoretically expected Laplace distribution of $p = 0.57$.
+In order to test our theoretical predictions for the distributions governing (averaged) cospectra, we first produced $3260$ individual FAD-corrected cospectra out of each of the $10\,\mathrm{s}$ segments, and for each cospectrum extracted the cospectral densities in the range $50\,\mathrm{Hz}$ to $400\,\mathrm{Hz}$, where we do not expect strong contamination by the pulsar signal and its harmonics, or by the low-frequency red noise component.
+We plot a histogram of cospectral densities between $50\,\mathrm{Hz}$ and $400\,\mathrm{Hz}$ for all $3260$ segments (a total of $11406500$ cospectral densities) in Figure \ref{fig:herx1_cs} (lower left panel), along with the theoretically expected Laplace distribution.
+We find a generally high agreement between the Laplace distribution and the distribution of observed cospectral densities.
 
@@
 As with the single cospectrum, the histogram for the averaged cospectral densities matches the theoretical prediction very closely.
-We compute the two-sided Kolmogorov-Smirnov statistic for the histogram of averaged cospectral densities, compared to the theoretical expected distribution, as$k = 0.01$, equivalent to a p-value of $p = 0.19$.
 This indicates that there is no evidence in the data that would argue against the observed cospectral densities being drawn from the distribution in Equation \ref{eqn:averaged_cdf}.

### 2017-12-19 paper/cospectra-paper.tex
Some minor wording changes and spelling fixes

@@
 We provide an instructive example of a quasi-periodic oscillation in \nustar\ and show that applying standard periodogram statistics leads to underestimated tail probabilities for period detection.
+\textbf{We also demonstrate the application of these distributions to a \nustar\ observation of the X-ray pulsar Hercules X-1.}
 
@@
 
-Within astronomy, and in particular the field of X-ray timing, it has recently been recognized \textbf{ as a powerful solution to two different problems.
+Within astronomy, and in particular the field of X-ray timing, it has recently been recognized \textbf{as a powerful solution to two different problems.
 One problem is the reliable estimation of the variability contained in certain features of the power spectrum like quasi-periodic oscillations (QPOs) via the fractional rms amplitude.
@@
 
-The cospectrum is also effectively used as as one approach to mitigate instrumental effects , in particular an effect called \textit{dead time}.
+The cospectrum is also effectively used as as one approach to mitigate instrumental effects, in particular an effect called \textit{dead time}.
 \textbf{In many X-ray detectors, after detection of a photon there is a time interval during which the detector (or in imaging detectors the individual pixel) is blocked from detecting a second photon.
@@
 \textbf{We note that the results below hold only for cases where the signal to be detected is phase-aligned in both light curves.
-This must be true by definition for analyses utilizing data from independent, identical detectors observing the same source simultaneously (unless detectors are so far apart that light travel time effects might be an issue and the data is not barycentred to the gravitational centre of the solar system), but might not strictly be true for observations of the same source e.g.\ in different energy bands.
+This must be true by definition for analyses utilizing data from independent, identical detectors observing the same source simultaneously (unless detectors are so far apart that light travel time effects might be an issue and photon arrival times are not barycentre-corrected to the gravitational centre of the solar system), but might not strictly be true for observations of the same source e.g.\ in different energy bands.
 We caution the reader that for signals that are not phase-aligned, the resulting power may decrease significantly and result in a non-detectable signal.}
@@
 We show that the statistical distribution indeed changes from a Laplace distribution once multiple cospectra are averaged and becomes consecutively more Gaussian as a larger number of cospectra are included in the average in Section \ref{sec:averaged_detthres}.
-Finally, Section \ref{sec:nustarqpo} presents a real-world example using simulated \nustar\ data of a quasi-periodic oscillation (QPO) as commonly found in accreting neutron star X-ray binaries.
+Finally, Section \ref{sec:nustarqpo} presents two real-world examples: the first uses simulated \nustar\ data of a quasi-periodic oscillation (QPO) as commonly found in accreting neutron star X-ray binaries.
+The second example comprises a real \nustar\ observation of the bright X-ray pulsar Hercules X-1.
 We end in Section \ref{sec:discussion} with a short discussion and conclusion.
@@
 We assume that both the white noise and the periodic signal are wide-sense stationary (i.e.
-the mean and the autocovariance of the time series do not change with time) and the light curve is evenly sampled simultaneously in two identical detectors.
+the mean and the autocovariance of the time series do not change with time) and the light curve is evenly sampled simultaneously in two identical detectors, or that photon arrival times are binned in intervals of equal length.
 While the white noise is also assumed to be strictly stochastic, the signal to be detected may be either deterministic (in the case of a strictly periodic signal) or stochastic (for a quasi-periodic oscillation).
@@
 \mathcal{F}_x(j) &= & \frac{1}{2} (A_{xj} - i B_{xj}) e^{-i\left( \frac{2 \pi j t}{T} \right)} \\
-\mathcal{F}_y(j) &= & \frac{1}{2} (A_{yj} - i B_{yj}) e^{-i\left( \frac{2 \pi j t}{T} \right)} \, ,
+\mathcal{F}_y(j) &= & \frac{1}{2} (A_{yj} - i B_{yj}) e^{-i\left( \frac{2 \pi j t}{T} \right)} \, .
 \end{eqnarray}
 
-\noindent where $i = \sqrt{-1}$, and $A_{xj}, A_{yj}$ and $B_{xj}, B_{yj}$ describe the real and imaginary parts of the Fourier amplitudes, respectively (for a pedagogical introduction into Fourier analysis, see \citealt{vanderklis1989}).
+\noindent Here, $i = \sqrt{-1}$, and $A_{xj}, A_{yj}$ and $B_{xj}, B_{yj}$ describe the real and imaginary parts of the Fourier amplitudes, respectively (for a pedagogical introduction into Fourier analysis, see \citealt{vanderklis1989}).
 \textbf{We restrict $\mathcal{F}_x(j)$ and $\mathcal{F}_y(j)$ to frequencies between $\nu_{j=0} = 1/T$ and the Nyquist frequency $\nu_{j=N/2} = 1/(2\Delta t)$.}
@@
 \end{equation}
-but since $\mu_x = \mu_y = 0$ for the Fourier amplitudes \textbf{of stationary stochastic processes}, this reduces to
+but since $\mu_x = \mu_y = 0$ for the Fourier amplitudes \textbf{of a stationary stochastic process}, this reduces to
 \begin{equation}
@@
 \textbf{Note that for light curves affected by dead time, the resulting cospectrum will still follow the Laplace distribution above, but with a variable variance that changes as a function of frequency \citep{Bachetti+15}.
-In practice, this issue can be straightforwardly corrected using the differences in Fourier amplitudes derived from the light curves of two detectors, and \citet{bachetti2017} show that the corrected cospectrum will closely follow the Laplace distribution derived here, allowing for unbiased significance tests for periodicity detection.}
+In practice, the cospectrum can be straightforwardly corrected for this effect using the differences in Fourier amplitudes derived from the light curves of two detectors (the \textit{Fourier Amplitude Difference} (FAD) technique), and \citet{bachetti2017} show that the corrected cospectrum will closely follow the Laplace distribution derived here, allowing for unbiased significance tests for periodicity detection.}
 
@@
 
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{qpo.png}
+\caption{Comparison of the averaged periodograms from the two detectors (blue), the periodogram obtained by the sum of the light curves (yellow; both left panel) and the cospectrum (right panel), for $3000\,\mathrm{s}$ of synthetic \nustar data with an incident count rate of $200 \mathrm{counts}\mathrm{s}^{-1}$ and a strong QPO at $200\,\mathrm{Hz}$.
+The QPO has an rms of $\sim$9\% and a high Q factor of $\sim$40.
+The shape of the periodogram is strongly distorted from the expected flat power spectrum centred on a constant value of $2$, with deviations of more that $0.5$ in the mean level of the densities common.
+In dashed lines, we also show the expected (trial-corrected) $99\%$ confidence level for the periodograms (in the same colour, left panel) as well as the cospectrum (red, right panel).
+The signal is detected in both the periodogram of the combined light curves as well as the cospectrum.
+However, because the white noise level is variable in the periodogram and the power spectrum dips below the expected level of $2$ around the frequency of the QPO, the signal is not detected as significantly as it should have been if no dead time was present.
+Assessing the significance is much more straightforward in the cospectrum, which retains a flat baseline.
+}
+\label{fig:qpo}
+\end{center}
+\end{figure*}
 \section{\textbf{Caveats}}
@@
 \textbf{An additional assumption is the phase-alignment of the periodic signal to be detected in both light curves.
-This is always true when the same source is observed simultaneously in the same energy range in independent, but identical detectors on the same spacecraft, as is common practice for instruments like \nustar\, \astrosat\ and \rxte.
+This is always true when the same source is observed simultaneously in the same energy range in independent, but identical detectors on the same spacecraft, as is common practice for instruments like \nustar, \astrosat\ and \rxte.
 Because the phase of the periodic signal can more generally be energy-dependent, we caution the reader to be careful when producing cospectra of light curves taken for example in different energy ranges.
@@
 
-\textbf{In realistic applications, in particularly where dead time is a major concern, the cospectrum will show these detector effects in the form of a frequency-dependent variation in the local variance.
+\textbf{In realistic applications, in particularly where dead time is a major concern, the cospectrum will show a frequency-dependent variation in the local variance, whose strength depends positively on the overall count rate of the object observed.
 It is imperative that this should be corrected before using the distributions used here by applying the Fourier Amplitude Differencing (FAD) technique \citep{bachetti2017}.
@@
 %\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{qpo.png}
-\caption{Comparison of the averaged periodograms from the two detectors (blue), the periodogram obtained by the sum of the light curves (yellow; both left panel) and the cospectrum (right panel), for $3000\,\mathrm{s}$ of synthetic \nustar data with an incident count rate of $200 \mathrm{counts}\mathrm{s}^{-1}$ and a strong QPO at $200\,\mathrm{Hz}$.
-The QPO has an rms of $\sim$9\% and a high Q factor of $\sim$40.
-The shape of the periodogram is strongly distorted from the expected flat power spectrum centred on a constant value of $2$, with deviations of more that $0.5$ in the mean level of the densities common.
-In dashed lines, we also show the expected (trial-corrected) $99\%$ confidence level for the periodograms (in the same colour, left panel) as well as the cospectrum (red, right panel).
-The signal is detected in both the periodogram of the combined light curves as well as the cospectrum.
-However, because the white noise level is variable in the periodogram and the power spectrum dips below the expected level of $2$ around the frequency of the QPO, the signal is not detected as significantly as it should have been if no dead time was present.
-Assessing the significance is much more straightforward in the cospectrum, which retains a flat baseline.
-}
-\label{fig:qpo}
-\end{center}
-\end{figure*}
 
@@
 
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{nustar_lc_psd.pdf}
-\caption{\textbf{Upper panel: Light curve of Her X-1 observed with \nustar, observation ID 30002006002.
-The observation totals $32.67\,\mathrm{ks}$, the light curve presented here is produced by combining event data from both FPMA and FPMB detectors.
-Gaps in the light curve are caused by the $96.8 \,\mathrm{m}$ orbit around Earth.
-Lower panel: averaged periodogram of $3260$ $10\,\mathrm{s}$-duration segments.
-The bright pulsar produces a very strong periodic signal at $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$ ($P_\mathrm{rot} = 1.24\,\mathrm{s}$) and at the signal's harmonics.
-In order to show the effects of dead time on the underlying detector noise, we have zoomed into the noise region, such that the peaks produced by the pulsar extend beyond the upper edge of the plot.
-At the lowest frequencies, the periodogram deviates from the expected $\chi^2$ distribution centered on $\mu = 2$ because of longer-term, aperiodic variability in the source itself.
-At frequencies above a $\sim$5 Hz or so, dead time strongly modulates the noise powers into an oscillatory pattern, indicating that the use of standard statistical distributions commonly used to find periodic or quasi-periodic signals will lead to biased results.
-}}
-\label{fig:herx1_lc}
-\end{center}
-\end{figure*}
-
 \subsection{A Simulated QPO in \nustar}
@@
 The baseline of the cospectrum, conversely, is not distorted by dead time, and requires only an estimate of the \textit{local} rms in order to calculate the significance of the QPO using Equation~\ref{eqn:averaged_cdf}: it is sufficient to \textit{multiply the cospectrum around the feature by an estimate of the local standard deviation of the white noise} (which is 1 in the ideal case) to use the equations above with no modifications.
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{nustar_lc_psd.pdf}
+\caption{\textbf{Upper panel: Light curve of Her X-1 observed with \nustar, observation ID 30002006002.
+The observation totals $32.67\,\mathrm{ks}$, the light curve presented here is produced by combining event data from both FPMA and FPMB detectors.
+Gaps in the light curve are caused by the $96.8 \,\mathrm{m}$ orbit around Earth.
+Lower panel: averaged periodogram of $3260$ $10\,\mathrm{s}$-duration segments.
+The bright pulsar produces a very strong periodic signal at $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$ ($P_\mathrm{rot} = 1.24\,\mathrm{s}$) and at the signal's harmonics.
+In order to show the effects of dead time on the underlying detector noise, we have zoomed into the noise region, such that the peaks produced by the pulsar extend beyond the upper edge of the plot.
+At the lowest frequencies, the periodogram deviates from the expected $\chi^2$ distribution centered on $\mu = 2$ because of longer-term, aperiodic variability in the source itself.
+At frequencies above a $\sim$5 Hz or so, dead time strongly modulates the noise powers into an oscillatory pattern, indicating that the use of standard statistical distributions commonly used to find periodic or quasi-periodic signals will lead to biased results.
+}}
+\label{fig:herx1_lc}
+\end{center}
+\end{figure*}
 
@@
 
-Her X-1 was observed by the Nuclear Spectroscopic Telescope Array (\nustar) multiple times in the last years. 
+Her X-1 has been observed by the Nuclear Spectroscopic Telescope Array (\nustar) multiple times. 
 For this work, we considered the observation taken from UT 2012-09-19 to UT 2012-09-20, one of those used by \citet{fuerst13} to characterize the cyclotron resonance scattering features in the spectrum of the source.
@@
 
-As an example, we have simulated how both a bright constant light curve and a QPO would appear in \nustar\ in the presence of dead time, and have shown that the shape of the periodogram is strongly distorted, whereas that of the cospectrum is not (for a longer introduction into the cospectrum and how it can be used in the presence of dead time, see also \citealt{Bachetti+15}).
+As an example, we have simulated how a QPO would appear in \nustar\ in the presence of dead time, and have shown that the shape of the periodogram is strongly distorted, whereas that of the cospectrum is not (for a longer introduction into the cospectrum and how it can be used in the presence of dead time, see also \citealt{Bachetti+15}).
 The significance of the QPO is difficult to assess in the periodogram, because of the non-linearities introduced by the variable dead time.
@@
 The cospectrum, on the other hand, only requires an estimate of the local variance in order to use the equations derived above, making it a far more convenient choice for periodicity detection.
-\textbf{Using a straightforward correction using the differences between Fourier amplitudes derived from two independent detectors, cospectra can be corrected such that they follow the distributions derived here very closely.}
+\textbf{By applying a straightforward correction using the differences between Fourier amplitudes derived from two independent detectors, cospectra can be corrected such that they follow the distributions derived here very closely.}
 
@@
 The distributions laid out above allow for detecting periodic and narrow quasi-periodic signals in the presence of detector white noise, and especially important in the context of pulsar detection in X-rays, where faint sources may yield marginal detections even in the best of cases.
-At the same time, as instruments like \astrosat\ and \ixpe, \textbf{with similar set-ups involving multiple identical detectors observing the same source simultaneously,} allow for observations with higher sensitivity, incorporating an accurate treatment of instrumental biases becomes increasingly important, and the cospectral statistics laid out here provide powerful tools to do so.
+At the same time, as instruments like \astrosat\ and \ixpe\ allow for observations with higher sensitivity, incorporating an accurate treatment of instrumental biases becomes increasingly important, and the cospectral statistics laid out here provide powerful tools to do so.
+
 Notably absent from this discussion, however, is the much more common case where a source exhibits stochastic variability in the form of red noise or notably broadened quasi-periodic oscillations.

### 2017-12-19 paper/referee_response.md
Minor changes to referee response

@@
 We believe that some of the issues raised by the referee may be result of a misunderstanding, and have realized that our text was not as clear as it should have been.
-We have expanded explanations in several cases, and have added a new section with caveats, as well as a data analysis on real NuSTAR data for illustrative purposes (all changes are marked in bold).
+We have expanded explanations in several cases, and have added a new section with caveats, as well as a data analysis on real NuSTAR data for illustrative purposes (all changes are marked in bold except where clearly stated).
 Below, we lay out in detail our changes to the manuscript, and respond directly to the points the referee has raised.
 We hope that this helps solve the misunderstanding and satisfactorily addresses the referee's concern.
+
+One small change not marked in bold (to avoid confusion with mathematical notation) was a typo in the definition of D directly following Equation 19, where a factor of j! was missing.
 
@@
 The practice of using the cospectrum of time series observed from two independent detectors, rather than the periodogram of the combined time series, was pioneered in NuSTAR as a means to mitigate the effects of dead time (and has since been used is >20 papers).
-However, the *underlying light curve* from the source is exactly the same, and only modified by detectore effects and Poisson counting statistics.
+However, the *underlying light curve* from the source is exactly the same, and only modified by detector effects and Poisson counting statistics.
 By definition, any periodic signal, be it in an X-ray pulsar or an accreting X-ray binary, must be strictly phase-aligned if it was observed simultaneously with detectors on the same spacecraft.
@@
 
-As described above, the common practice of using the cospectrum derived from two identical detectors implies that we disagree with the referee on the point that the case discussed in the paper does not exist in reality.
-In particular, the primary reason this manuscript came into being is that we were both approached independently and asked how to define the statistical distributions for the cospectrum in the case of X-ray pulsation searches with NuSTAR.
+As described above, it is now common practice to use the cospectrum derived from two identical detectors to search for variability.
+As an aside, the primary reason this manuscript came into being is that we were both approached independently and asked how to define the statistical distributions for the cospectrum in the case of X-ray pulsation searches with NuSTAR.
 This, combined with the erroneous assumptions made about the statistical distributions used in the current astronomy literature on the topic (primarily the use of the Chi-square and normal distributions), led us to believe that a manuscript laying out the correct statistical distributions to be used is warranted.
@@
 The resulting noise distributions are narrow, but deviate very strongly from the expected Chi-square distribution with a mean of 2 and a variance of 2/sqrt(M), where M=3260.
-This is the primary motivation for the cospectrum as an alternative measure unaffected by dead time.
+This is one of the two primary motivations for the cospectrum as an alternative measure unaffected by dead time.
 

### 2018-01-08 paper/cospectra-paper.tex
Changes in response to referee report and new version of referee response

@@
 In recent years, the cross spectrum has received considerable attention as a means of characterising the variability of astronomical sources as a function of wavelength.
-While much has been written about the statistics of time and phase lags, the cospectrum has only recently been understood as means of mitigating instrumental effects dependent on temporal frequency in astronomical detectors, as well as a method of characterizing the coherent variability in two  wavelength ranges on different time scales.
-In this paper, we lay out the statistical foundations of the cospectrum, starting with the simplest case of detecting a periodic signal in the presence of white noise.
-This case is especially relevant for detecting faint X-ray pulsars in detectors heavily affected by instrumental effects, including \nustar, \astrosat\ and \ixpe.
-We show that the statistical distributions of both single and averaged cospectra differ considerably from those for standard periodograms.
+The cospectrum has only recently been understood as means of mitigating instrumental effects dependent on temporal frequency in astronomical detectors, as well as a method of characterizing the coherent variability in two wavelength ranges on different time scales.
+In this paper, we lay out the statistical foundations of the cospectrum, starting with the simplest case of detecting a periodic signal in the presence of white noise, \textbf{under the assumption that the same source is observed simultaneously in independent detectors in the same energy range.} This case is especially relevant for detecting faint X-ray pulsars in detectors heavily affected by instrumental effects, including \nustar, \astrosat\ and \ixpe, \textbf{which allow for even sampling and where the cospectrum can act as an effective way to mitigate dead time.} We show that the statistical distributions of both single and averaged cospectra differ considerably from those for standard periodograms.
 While a single cospectrum follows a Laplace distribution exactly, averaged cospectra are approximated by a Gaussian distribution only for more than $\sim\!\!
@@
 We provide an instructive example of a quasi-periodic oscillation in \nustar\ and show that applying standard periodogram statistics leads to underestimated tail probabilities for period detection.
-\textbf{We also demonstrate the application of these distributions to a \nustar\ observation of the X-ray pulsar Hercules X-1.}
+We also demonstrate the application of these distributions to a \nustar\ observation of the X-ray pulsar Hercules X-1.
 
@@
 Fourier analysis, especially through the periodogram\footnote{We distinguish in this paper between the \textit{power spectrum}, which describes the process at the source generating variable time series, and the \textit{periodogram}, which denotes a realization of said power spectrum, i.e.\ the time series we actually observe.
-\textbf{In line with the signal processing literature, we also use the term \textit{periodogram} or \textit{power spectral density} for the square of the absolute value of Fourier amplitudes derived from an observed light curve.
-We use the terms \textit{cospectrum} or \textit{cospectral density} to denote the real component of the cross-spectrum, i.e.\ the result of a multiplication of the Fourier amplitudes of one light curve with the complex conjugate of the Fourier amplitudes of a second light curve (see also Section \ref{sec:whitenoise_cospectra} for an exact definition)}}, has long been used to find periodic and quasi-periodic signals as well as characterize the stationary stochastic processes often present in accreting systems.
+In line with the signal processing literature, we also use the term \textit{periodogram} or \textit{power spectral density} for the square of the absolute value of Fourier amplitudes derived from an observed light curve.
+We use the terms \textit{cospectrum} or \textit{cospectral density} to denote the real component of the cross-spectrum, i.e.\ the result of a multiplication of the Fourier amplitudes of one light curve with the complex conjugate of the Fourier amplitudes of a second light curve (see also Section \ref{sec:whitenoise_cospectra} for an exact definition)}, has long been used to find periodic and quasi-periodic signals as well as characterize the stationary stochastic processes often present in accreting systems.
 While in principle, the statistics of the periodogram is well understood and characterized in the literature \citep[e.g.][]{vanderklis1989}, the periodogram is often subject to instrumental effects like dead time that change its statistical properties and thus make statistical inference difficult in practice \citep[e.g.][]{Zhang+95}.
 
-In the past years, the field of \textit{spectral timing} has enjoyed significant success by making it possible to combine both temporal and spectral information in a single model.
-Within this framework, the complex cross spectrum---defined as the Fourier transform of one time series with the complex conjugate of the Fourier transform of a second time series--holds a central position.
-The cross spectrum is commonly used to compute phase lags, which are easily converted to time lags and are central for understanding e.g.\ reverberation mapping in accreting black holes \citep[see][for a recent review]{uttley2014}.
+The real part of the cross spectrum---defined as the Fourier transform of one time series with the complex conjugate of the Fourier transform of a second time series--,also named \textit{cospectrum} has long been used to study gravity waves in the Earth's atmosphere \citep[e.g.][]{john2016}, models of the Martian atmosphere \citep[e.g.][]{wang2016}, the solar heliosphere \citep[e.g.][]{vigeesh2017}, surface elevation of arctic sea ice \citep[e.g.][]{ardhuin2016} and drifting snow \citep[e.g.][]{paterna2016} as well as surface gravity waves in beaches \citep[e.g.][]{fiedler2015} and eddy heat flux in the Earth's troposphere \citep[e.g.][]{wang2015,zurita-gotor2017}.
 
-The real part of the cross spectrum, also named \textit{cospectrum}, has gained less attention, but can be just as useful.
-It has long been used to study gravity waves in the Earth's atmosphere \citep[e.g.][]{john2016}, models of the Martian atmosphere \citep[e.g.][]{wang2016}, the solar heliosphere \citep[e.g.][]{vigeesh2017}, surface elevation of arctic sea ice \citep[e.g.][]{ardhuin2016} and drifting snow \citep[e.g.][]{paterna2016} as well as surface gravity waves in beaches \citep[e.g.][]{fiedler2015} and eddy heat flux in the Earth's troposphere \citep[e.g.][]{wang2015,zurita-gotor2017}.
-
-Within astronomy, and in particular the field of X-ray timing, it has recently been recognized \textbf{as a powerful solution to two different problems.
+Within astronomy, and in particular the field of X-ray timing, the cospectrum has recently been recognized as a powerful solution to two different problems.
 One problem is the reliable estimation of the variability contained in certain features of the power spectrum like quasi-periodic oscillations (QPOs) via the fractional rms amplitude.
@@
 For instruments with two independent, identical detectors, the cospectrum of the light curves measured in each detector, respectively, will not contain any white noise contribution, since Poisson counting statistics are effects local to the instrument, and thus the observed data sets are independently sampled in each detector.
-This makes estimation of source-intrinsic variability much more reliable than comparable measurements using the periodogram.}
+This makes estimation of source-intrinsic variability much more reliable than comparable measurements using the periodogram.
 
 The cospectrum is also effectively used as as one approach to mitigate instrumental effects, in particular an effect called \textit{dead time}.
-\textbf{In many X-ray detectors, after detection of a photon there is a time interval during which the detector (or in imaging detectors the individual pixel) is blocked from detecting a second photon.
-This time interval is generally called \textit{dead time} (the detector is effectively ``dead'', and incoming photons will not produce a signal) and its characteristic time scale is set by the details of the detector and can range from very short ($2\mu\mathrm{s}$ in the \fermi\ Gamma-Ray Burst Monitor) to very long ($2.5\mathrm{ms}$ in \nustar).} It leads to frequency-dependent changes in the mean and variance of the statistical distributions governing the periodogram, \textbf{which cannot be mitigated by averaging periodograms of multiple segments}.
+In many X-ray detectors, after detection of a photon there is a time interval during which the detector (or in imaging detectors the individual pixel) is blocked from detecting a second photon.
+This time interval is generally called \textit{dead time} (the detector is effectively ``dead'', and incoming photons will not produce a signal) and its characteristic time scale is set by the details of the detector and can range from very short ($2\mu\mathrm{s}$ in the \fermi\ Gamma-Ray Burst Monitor) to very long ($2.5\mathrm{ms}$ in \nustar).
+It leads to frequency-dependent changes in the mean and variance of the statistical distributions governing the periodogram, which cannot be mitigated by averaging periodograms of multiple segments.
 
-Where in standard analysis, light curves of multiple detectors are summed before Fourier transforming the summed light curve, it is possible to instead Fourier-transform the signal of two independent detectors \textbf{within the same instrument observing the same source strictly simultaneously in the same energy band}.
-The resulting \textbf{co-spectral} densities will be less affected by dead time (see details in \citealt{Bachetti+15}), because the latter is introduced in each detector independently and tends to cancel out in the cospectrum.
-This approach relevant for current X-ray missions \textbf{carrying at least two identical detectors} such as \nustar\ and \astrosat, but will also be relevant to future missions with multiple detectors like \ixpe.
-It has recently been used in \nustar\ studies of millisecond pulsars \citep{ferrigno2017}, Ultraluminous X-ray Sources \citep{bachetti2016} and X-ray binaries \citep{barriere2015,zoghbi2016,ingram2016,huppenkothen2017,stiele2017}.
-\textbf{It is particularly important for instruments like \nustar\ which have a fairly long dead time and thus exhibit dead time effects in the periodogram at frequencies comparable to those where signals are expected, and at fairly low count rates}.
-Similarly, the cospectrum of two time series taken with the same instrument, but in different wavelength ranges, can be used to characterize the coherent, \textbf{phase-aligned} variability in both time series as a function of frequency.
-
-While much has been written on the subject of the statistics of cross spectra and time lags \citep[e.g.][]{epitropakis2016}, the derivation of cospectral statistics is notably absent from the astronomy literature, and most publications \textbf{utilizing the approach of forming co-spectra from identical, but independent detectors observing the same source} assume that either the $\chi^2_2$ distribution used for standard periodograms or a Gaussian distribution for averaged spectra is appropriate.
+While much has been written on the subject of the statistics of cross spectra and time lags \citep[e.g.][]{epitropakis2016}, the derivation of cospectral statistics is notably absent from the astronomy literature, and most publications utilizing the approach of forming co-spectra from identical, but independent detectors observing the same source assume that either the $\chi^2_2$ distribution used for standard periodograms or a Gaussian distribution for averaged spectra is appropriate.
 That these assumptions are valid has not been shown until now.
@@
 In this paper, we lay out the basic statistical distributions for detecting periodic and narrow quasi-periodic signals in the presence of detector noise (e.g.\ photon counting statistics) for both single and averaged cospectra.
-We show below that unlike for the periodogram, the statistical distribution for a single cospectrum reduces to a Laplace distribution, while the distribution for averaged cospectra is considerably more complex.
+\textbf{The primary use case for the distributions defined in the manuscript are cases where observations were taken simultaneously of the same source with independent, identical detectors, such that the source light curve is the same, modified only by detector effects and Poisson counting statistics.
+This is now standard practice for multiple X-ray missions (including the now defunct \rxte, currently operating \nustar\ and \astrosat, and future missions like \ixpe).
+The practice of using the cospectrum of time series observed from two independent detectors, rather than the periodogram of the combined time series, was pioneered in \nustar\ as a means to mitigate the effects of dead time \citep{Bachetti+15}.
+It has recently been used in \nustar\ studies of millisecond pulsars \citep{ferrigno2017}, Ultraluminous X-ray Sources \citep{bachetti2016} and X-ray binaries \citep{barriere2015,zoghbi2016,ingram2016,huppenkothen2017,stiele2017}.}
+
+We show below that unlike for the periodogram, the statistical distribution for a single cospectrum \textbf{under the assumptions above} reduces to a Laplace distribution, while the distribution for averaged cospectra is considerably more complex.
 We also find that for averaged cospectra consisting of more than $\sim$30 averaged individual segments, the assumption of a Gaussian distribution is indeed appropriate for single-trial tail probabilities and reduces computation overhead.
 However, for averaged cospectra of fewer segments, more stringent significance thresholds or large numbers of trials, the statistical distribution--and hence the derived $p$-values used for period detection--deviate significantly from a Gaussian distribution.
-\textbf{We note that the results below hold only for cases where the signal to be detected is phase-aligned in both light curves.
+We note that the results below hold only for cases where the signal to be detected is phase-aligned in both light curves.
 This must be true by definition for analyses utilizing data from independent, identical detectors observing the same source simultaneously (unless detectors are so far apart that light travel time effects might be an issue and photon arrival times are not barycentre-corrected to the gravitational centre of the solar system), but might not strictly be true for observations of the same source e.g.\ in different energy bands.
-We caution the reader that for signals that are not phase-aligned, the resulting power may decrease significantly and result in a non-detectable signal.}
+We caution the reader that for signals that are not phase-aligned \textbf{(for example because energy-dependent phase shifts exist in the source)}, the resulting power may decrease significantly and result in a non-detectable signal.
 
@@
 
-For the reader looking for the statistical distributions of relevance, who may be only casually interested in the mathematical background, we point to Equations \ref{eqn:csdist} and \ref{eqn:averaged_pdf} for the probability density functions (PDFs) for a single cospectrum and averaged cospectrum, respectively, and Equations \ref{eqn:cospectrum_cdf} and \ref{eqn:averaged_cdf} for the cumulative distribution functions (CDFs) in both cases.
+We point readers looking for the relevant statistical distributions, who may be only casually interested in the mathematical background, to Equations \ref{eqn:csdist} and \ref{eqn:averaged_pdf} for the probability density functions (PDFs) for a single cospectrum and averaged cospectrum, respectively, and Equations \ref{eqn:cospectrum_cdf} and \ref{eqn:averaged_cdf} for the cumulative distribution functions (CDFs) in both cases.
 
@@
 
-\textbf{In the following, we will consider the example of detecting a strictly periodic or very narrow quasi-periodic signal in the presence of simple white noise, as is commonly the case for example in pulsar searches in X-ray data.
+In the following, we will consider the example of detecting a strictly periodic or very narrow quasi-periodic signal in the presence of simple white noise, as is commonly the case for example in pulsar searches in X-ray data.
 We assume that both the white noise and the periodic signal are wide-sense stationary (i.e.
@@
 While the white noise is also assumed to be strictly stochastic, the signal to be detected may be either deterministic (in the case of a strictly periodic signal) or stochastic (for a quasi-periodic oscillation).
-We then aim to reject a null hypothesis where the observed cospectral density at a given frequency can be explained by white noise alone.}
+We then aim to reject a null hypothesis where the observed cospectral density at a given frequency can be explained by white noise alone.
 
@@
 
-Consider two independently distributed, \textbf{evenly-sampled}, constant, stationary time series,
+Consider two independently distributed, evenly-sampled, constant, stationary time series,
 
@@
 
-\noindent with $N$ data points taken at simultaneous time intervals $\{t_k\}_{k=1}^N$ \textbf{with a constant time resolution $\Delta t$ and a total duration $T = N\Delta t$}.
+\noindent with $N$ data points taken at simultaneous time intervals $\{t_k\}_{k=1}^N$ with a constant time resolution $\Delta t$ and a total duration $T = N\Delta t$.
 Assume for simplicity that the measurements $x_k$ and $y_k$ are normally distributed, such that
@@
 \noindent Here, $i = \sqrt{-1}$, and $A_{xj}, A_{yj}$ and $B_{xj}, B_{yj}$ describe the real and imaginary parts of the Fourier amplitudes, respectively (for a pedagogical introduction into Fourier analysis, see \citealt{vanderklis1989}).
-\textbf{We restrict $\mathcal{F}_x(j)$ and $\mathcal{F}_y(j)$ to frequencies between $\nu_{j=0} = 1/T$ and the Nyquist frequency $\nu_{j=N/2} = 1/(2\Delta t)$.}
+We restrict $\mathcal{F}_x(j)$ and $\mathcal{F}_y(j)$ to frequencies between $\nu_{j=0} = 1/T$ and the Nyquist frequency $\nu_{j=N/2} = 1/(2\Delta t)$.
 The complex cross spectrum is then calculated by multiplying the Fourier transform of light curve $\mathbf{x}$ with the complex conjugate of the Fourier transform of light curve $\mathbf{y}$ (\citealt{vaughan1997,nowak1999}, see also \citealt{uttley2014} for a recent review of spectral timing techniques):
@@
 
-\noindent For \textbf{evenly sampled}, normally distributed light curves (and indeed for uncertainties coming from a wide range of statistical distributions, including the Poisson distribution, if $N$ is large), the real and imaginary amplitude components are distributed as $A_{xj}, B_{xj} \sim \mathcal{N}(0, \sigma_x^2)$ with $\sigma_x =  \sqrt{\frac{\sum_{k=1}^{N}{x_k}}{2}}$ and  $A_{yj}, B_{yj} \sim \mathcal{N}(0, \sigma_y^2)$ with $\sigma_y = \sqrt{\frac{\sum_{k=1}^{N}{y_k}}{2}}$.
-\textbf{Note that these distributions for the Fourier amplitudes only hold if the underlying process producing the observations is stochastic and wide-sense stationary.
+\noindent For evenly sampled, normally distributed light curves (and indeed for uncertainties coming from a wide range of statistical distributions, including the Poisson distribution, if $N$ is large), the real and imaginary amplitude components are distributed as $A_{xj}, B_{xj} \sim \mathcal{N}(0, \sigma_x^2)$ with $\sigma_x =  \sqrt{\frac{\sum_{k=1}^{N}{x_k}}{2}}$ and  $A_{yj}, B_{yj} \sim \mathcal{N}(0, \sigma_y^2)$ with $\sigma_y = \sqrt{\frac{\sum_{k=1}^{N}{y_k}}{2}}$.
+Note that these distributions for the Fourier amplitudes only hold if the underlying process producing the observations is stochastic and wide-sense stationary.
 This includes many processes commonly observed in astrophysical sources such as white noise observed from a constant background, as well as red noise processes such as shot noise and other (broken) power-law power spectra often seen in AGN, and quasi-periodic oscillations with stochastic variations in amplitude and period of the observed signal, commonly observed in black hole X-ray binaries.
-For strictly deterministic processes (e.g.\ strictly periodic variations), the distributions of the Fourier amplitudes will not be centered on $\mu = 0$, and the distributions below will not be correct.}
+For strictly deterministic processes (e.g.\ strictly periodic variations), the distributions of the Fourier amplitudes will not be centered on $\mu = 0$, and the distributions below will not be correct.
 
@@
 \begin{eqnarray}
-p_Z(z) & = & \int_{-\infty}^{+\infty}{p_X(x) p_Y(\frac{z}{x}) \frac{1}{|x|} dx} \\\nonumber
+p_Z(z) & = & \int_{-\infty}^{+\infty}{p_X(x) p_Y(\frac{z}{x}) \frac{1}{|x|} dx} \nonumber \\
 	   & = &  \int_{-\infty}^{+\infty}{\frac{1}{2 \pi \sigma_x \sigma_y} \exp{-\frac{x^2}{2\sigma_x^2}} \exp{-\frac{(z/x)^2}{2\sigma_y^2}} \frac{1}{|x|} dx} \label{eqn:pz} \, .
@@
 \end{equation}
-but since $\mu_x = \mu_y = 0$ for the Fourier amplitudes \textbf{of a stationary stochastic process}, this reduces to
+but since $\mu_x = \mu_y = 0$ for the Fourier amplitudes of a stationary stochastic process, this reduces to
 \begin{equation}
@@
 
-\textbf{Note that for light curves affected by dead time, the resulting cospectrum will still follow the Laplace distribution above, but with a variable variance that changes as a function of frequency \citep{Bachetti+15}.
-In practice, the cospectrum can be straightforwardly corrected for this effect using the differences in Fourier amplitudes derived from the light curves of two detectors (the \textit{Fourier Amplitude Difference} (FAD) technique), and \citet{bachetti2017} show that the corrected cospectrum will closely follow the Laplace distribution derived here, allowing for unbiased significance tests for periodicity detection.}
+Note that for light curves affected by dead time, the resulting cospectrum will still follow the Laplace distribution above, but with a variable variance that changes as a function of frequency \citep{Bachetti+15}.
+In practice, the cospectrum can be straightforwardly corrected for this effect using the differences in Fourier amplitudes derived from the light curves of two detectors (the \textit{Fourier Amplitude Difference} (FAD) technique), and \citet{bachetti2017} show that the corrected cospectrum will closely follow the Laplace distribution derived here, allowing for unbiased significance tests for periodicity detection.
 
@@
 \caption{Histogram of cospectral densities (grey) and the theoretical expectation (red) for three different cases.
-In the right panel, we show the distribution of densities for a single cospectrum, with its expected Laplace distribution from Equation \ref{eqn:laplace}, in the middle and right panel cospectral densities for averaging 10 (middle) and 100 (right) individual cospectra together.
+\textbf{We show 10000 cospectral densitiies in each panel in order to show the scatter around the theoretical distribution for small samples (see also Figure \ref{fig:herx1_cs} for a similar figure using real data and a much larger number of densities).} In the right panel, we show the distribution of densities for a single cospectrum, with its expected Laplace distribution from Equation \ref{eqn:laplace}, in the middle and right panel cospectral densities for averaging 10 (middle) and 100 (right) individual cospectra together.
 In the middle panel, the theoretical expectation of the sampling distribution is given by Equation \ref{eqn:averaged_pdf}, in the right panel by a Gaussian distribution with $\sigma=\sqrt{2/101}$.}
@@
 For practical purposes, evaluating this PDF for averaged spectra above $n\sim 85$ is difficult numerically, because the factorials and exponents in the sum become very large and small, respectively.
-However, as we will show in Section \ref{sec:averaged_detthres} below, \textbf{we expect that for large $n$, the Central Limit Theorem implies that the PDF of averaged cospectral densities tends towards a normal distribution.} We find that in practice, when $n \gtrsim 30$, detection thresholds derived from Equation \ref{eqn:averaged_pdf} provide only a negligible difference over that derived from a normal distribution $N(0, \sqrt{2/(n+1)}$, depending on the significance threshold required and the number of trials.
+However, as we will show in Section \ref{sec:averaged_detthres} below, we expect that for large $n$, the Central Limit Theorem implies that the PDF of averaged cospectral densities tends towards a normal distribution.
+We find that in practice, when $n \gtrsim 30$, detection thresholds derived from Equation \ref{eqn:averaged_pdf} provide only a negligible difference over that derived from a normal distribution $N(0, \sqrt{2/(n+1)}$, depending on the significance threshold required and the number of trials.
 
@@
 \end{figure*}
-\section{\textbf{Caveats}}
+\section{Caveats}
 
-\textbf{All caveats applying more generally to Fourier analysis and to periodograms specifically also apply to cospectra.
+All caveats applying more generally to Fourier analysis and to periodograms specifically also apply to cospectra.
 In particular, as for the $\chi^2$ distribution used for period searches in periodograms, the distributions derived here assume that the underlying process is wide-sense stationary and the light curves are simultaneously and evenly sampled.
 These conditions are typically met in most observations taken with X-ray timing instruments like \nustar\ and \rxte\ for problems such as timing of X-ray binaries, AGN and X-ray pulsar searches, but this may not be the case for e.g.\ optical observations of main-sequence or binary stars.
-Note that we also implicitly assume that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers, both of which may not be stationary.}
+Note that we also implicitly assume that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers, both of which may not be stationary.
 
-\textbf{Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}).
+Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}).
 Uneven sampling, on the other hand, introduces covariances between adjacent frequencies and renders many of the statistical assumptions underpinning period searches using both in the periodogram and the cospectrum invalid.
-For unevenly sampled data, other methods that do not depend on a regular sampling pattern must be employed (for example the Lomb-Scargle periodogram \citep{lomb1976, scargle1982} or the Bayesian period search in \citet{gregory1992}) and dead-time effects must be forward-simulated.}
+For unevenly sampled data, other methods that do not depend on a regular sampling pattern must be employed (for example the Lomb-Scargle periodogram \citep{lomb1976, scargle1982} or the Bayesian period search in \citet{gregory1992}) and dead-time effects must be forward-simulated.
 
-\textbf{An additional assumption is the phase-alignment of the periodic signal to be detected in both light curves.
+An additional assumption is the phase-alignment of the periodic signal to be detected in both light curves.
 This is always true when the same source is observed simultaneously in the same energy range in independent, but identical detectors on the same spacecraft, as is common practice for instruments like \nustar, \astrosat\ and \rxte.
@@
 In all these cases, period searches will only be sensitive to signals that are phase-aligned, and will in the worst case be undetectable if the phase-shift is 90 or 270 degrees.
-}
 
-\textbf{In realistic applications, in particularly where dead time is a major concern, the cospectrum will show a frequency-dependent variation in the local variance, whose strength depends positively on the overall count rate of the object observed.
+In realistic applications, in particularly where dead time is a major concern, the cospectrum will show a frequency-dependent variation in the local variance, whose strength depends positively on the overall count rate of the object observed.
 It is imperative that this should be corrected before using the distributions used here by applying the Fourier Amplitude Differencing (FAD) technique \citep{bachetti2017}.
@@
 In particularly, FAD-corrected cospectra tend to overestimate the integrated rms when both count rate and variability amplitude are both very high.
-This should be accounted for when deriving estimates of the fractional rms amplitude.}
-
-%\begin{figure*}
-%\begin{center}
-%\includegraphics[width=\textwidth]{whitenoise_deadtime.png}
-%\caption{\textbf{Left panel: Fourier products derived from white noise light curves heavily affected by dead time.
-There is a strong, frequency-dependent modulation in the periodogram (orange) introduced by dead time effects, rendering standard tests for periodicities unreliable.
-The cospectrum (dark grey) is less affected overall, but there remains a frequency-dependent variation in the local variance, with a narrower distribution governing the cospectral densities at low frequencies, and a wider distribution e.g.\ between 200 and 400 Hz.
-Applying the FAD correction (blue) produces a cospectrum that follows the distributions defined in this paper very closely, allowing for reliable periodicity testing.
-Right panel: histograms of cospectral densities derived from $1000$ simulated, dead time-affected cospectra (dark grey) and their FAD-corrected equivalents (blue).
-The dead time-affected cospectra overall have a narrower distribution (but note that this is integrated over all frequencies, thus the frequency dependence is washed out), whereas the corrected cospectra follow the expected distribution (red).
-%}}
-%\label{fig:wn_dt}
-%\end{center}
-%\end{figure*}
+This should be accounted for when deriving estimates of the fractional rms amplitude.
 
@@
 
-\textbf{In the following, we will consider two more realistic examples in turn.
+In the following, we will consider two more realistic examples in turn.
 First, we will consider a more realistic simulation of a QPO as expected to be observed in a typical \nustar\ observation.
-Subsequently, we present an example of real \nustar\ data containing a coherent pulsed signal from the X-ray neutron star source Hercules X-1.}
+Subsequently, we present an example of real \nustar\ data containing a coherent pulsed signal from the X-ray neutron star source Hercules X-1.
 
-%\subsection{Simulated Light Curves With Dead Time}
-
-%\begin{bf}
-%Here, we simulate simple white noise light curves and apply a typical \nustar-like dead time of $2.5\mathrm{ms}$ in order to explore how the distributions defined above work under more realistic conditions.
-We use the relatively extreme case of the Crab pulsar, one of the brightest variable X-ray sources in the sky [REF].
-We simulate event files for an observation of $T = 100\,\mathrm{s}$ and a constant count rate of $c = 400 \mathrm{counts}/\mathrm{s}$.
-Incident photon events are generated by picking $N=cT = 4 \times 10^{4}$ times of arrivals randomly from a uniform distribution between the start and end time, independently for each of the two \nustar\ detectors.
-We then apply dead time to both sets of photon events by excluding all photons that arrive within $2.5\mathrm{ms}$ of the previous photon arrival in each detector.
-For a count rate of $400\mathrm{counts}/\mathrm{s}$ and a \nustar-like dead time, this corresponds to a photon loss of $\sim$50\%.
-The resulting light curves are binned at a time resolution of $9.766 \times 10^{-4}\,\mathrm{s}$ (corresponding to a Nyquist frequency of $\nu_{\mathrm{Nyquist}} = 512 \,\mathrm{Hz}$) and averaged periodograms of each individual light curve, of the summed light curves, and the averaged cospectrum are created using segments of $t_\mathrm{seg} = 10\,\mathrm{s}$.
-We repeated this process 1000 times in order to gain sufficient simulated cospectral densities for statistical comparisons.
-
-%In Figure \ref{fig:wn_dt} (left panel), we show the periodogram, cospectrum and the corresponding probability distribution of cospectral densities of these simulations.
-In the periodogram, we see a large frequency-dependent deviation from the expected white noise level of $2$, which is difficult to correct for in practice and introduces a large frequency-dependent bias into any standard tests for periodicities that utilize the periodogram.
-The cospectrum, on the other hand, does not suffer from distortions of the underlying power spectrum, but does show frequency-dependent variations in the local variance (see also \citet{Bachetti+15} for a more in-depth explanation of this effect).
-%This indicates that the distributions defined in this paper cannot be applied to dead-time affected data without correction.
-Here, we follow the suggestion of \citet{bachetti2017} and use so-called \textit{Fourier Amplitude Differences} (FAD) in order to correct the cospectrum.
-In short, this technique rests on the idea that while variability intrinsic to the source will produce Fourier amplitudes with the same real and imaginary components, both dead time and Photon counting noise are independent between detectors, and will thus be uncorrelated between Fourier amplitudes derived from two identical, independent detectors.
-The \textit{difference} between Fourier amplitudes of one detector and another will hence effectively remove source-intrinsic variability and only contain dead-time related effects.
-One can then model the latter in the Fourier amplitude differences, and use this model to correct the original cospectrum (see \citet{bachetti2017} for details).
-
-%We corrected the cospectra using the FAD method and subsequently compared the empirical distributions derived from the simulated cospectra to the theoretical distribution defined in Section \ref{sec:averaged_detthres} (Figure \ref{fig:wn_dt}, right panel).
-We find that the FAD-corrected cospectra follow the derived distributions very closely even in this case involving an exceptionally high degree of dead time, and conclude that for most realistic scenarios, the distributions arrived here can be used to construct unbiased estimators for periodicities, assuming that the FAD correction is applied to the cospectra first.
-
-%\end{bf}
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{nustar_lc_psd.pdf}
+\caption{Upper panel: Light curve of Her X-1 observed with \nustar, observation ID 30002006002.
+The observation totals $32.67\,\mathrm{ks}$, the light curve presented here is produced by combining event data from both FPMA and FPMB detectors.
+Gaps in the light curve are caused by the $96.8 \,\mathrm{m}$ orbit around Earth.
+Lower panel: averaged periodogram of $3260$ $10\,\mathrm{s}$-duration segments.
+The bright pulsar produces a very strong periodic signal at $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$ ($P_\mathrm{rot} = 1.24\,\mathrm{s}$) and at the signal's harmonics.
+In order to show the effects of dead time on the underlying detector noise, we have zoomed into the noise region, such that the peaks produced by the pulsar extend beyond the upper edge of the plot.
+At the lowest frequencies, the periodogram deviates from the expected $\chi^2$ distribution centered on $\mu = 2$ because of longer-term, aperiodic variability in the source itself.
+At frequencies above a $\sim$5 Hz or so, dead time strongly modulates the noise powers into an oscillatory pattern, indicating that the use of standard statistical distributions commonly used to find periodic or quasi-periodic signals will lead to biased results.
+}
+\label{fig:herx1_lc}
+\end{center}
+\end{figure*}
 
@@
 Note that in all three cases we produced averaged periodograms and co-spectra by splitting the light curves into 600 segments of $5\,\mathrm{s}$ length each in order to suppress the variance in the powers and show the effects of dead time more clearly.
-\textbf{As mentioned above, cospectra with dead time are subject to frequency-dependent changes in variance, and we thus corrected the simulated cospectra using the FAD technique.}
+As mentioned above, cospectra with dead time are subject to frequency-dependent changes in variance, and we thus corrected the simulated cospectra using the FAD technique.
 
@@
 The baseline of the cospectrum, conversely, is not distorted by dead time, and requires only an estimate of the \textit{local} rms in order to calculate the significance of the QPO using Equation~\ref{eqn:averaged_cdf}: it is sufficient to \textit{multiply the cospectrum around the feature by an estimate of the local standard deviation of the white noise} (which is 1 in the ideal case) to use the equations above with no modifications.
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{nustar_lc_psd.pdf}
-\caption{\textbf{Upper panel: Light curve of Her X-1 observed with \nustar, observation ID 30002006002.
-The observation totals $32.67\,\mathrm{ks}$, the light curve presented here is produced by combining event data from both FPMA and FPMB detectors.
-Gaps in the light curve are caused by the $96.8 \,\mathrm{m}$ orbit around Earth.
-Lower panel: averaged periodogram of $3260$ $10\,\mathrm{s}$-duration segments.
-The bright pulsar produces a very strong periodic signal at $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$ ($P_\mathrm{rot} = 1.24\,\mathrm{s}$) and at the signal's harmonics.
-In order to show the effects of dead time on the underlying detector noise, we have zoomed into the noise region, such that the peaks produced by the pulsar extend beyond the upper edge of the plot.
-At the lowest frequencies, the periodogram deviates from the expected $\chi^2$ distribution centered on $\mu = 2$ because of longer-term, aperiodic variability in the source itself.
-At frequencies above a $\sim$5 Hz or so, dead time strongly modulates the noise powers into an oscillatory pattern, indicating that the use of standard statistical distributions commonly used to find periodic or quasi-periodic signals will lead to biased results.
-}}
-\label{fig:herx1_lc}
-\end{center}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{nustar_cospectra.pdf}
-\caption{\textbf{Upper panel: Cospectrum of a $10\,\mathrm{s}$ segment of the \nustar\ observations of Her X-1 (black).
-The cospectrum is created by observing the same source simultaneously, but statistically independently with the two identical FPMA and FPMB detectors.
-The rotational signal of the pulsar appears very clearly at $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$, but unlike the periodogram, the cospectrum is not affected by dead time.
-In red, we show the $p = 10^{-3}$ detection threshold, corrected for $6999$ trials (equivalent to the number of frequencies in the periodogram).
-Lower left:  histogram (grey) of the cospectral densities collected from 3260 individual cospectra between $50\,\mathrm{Hz}$ and $400\,\mathrm{Hz}$.
-In red, we also show the theoretically predicted Laplace distribution.
-Lower right: in black, the histogram of cospectral densities collected from $7$ averaged cospectra, each created from $15$ light curve segments of $200\,\mathrm{s}$ duration.
-Again, only cospectral densities between $50\,\mathrm{Hz}$ and $400\,\mathrm{Hz}$ were used in this figure.
-In red, we show the theoretical probability density function defined in Equation \ref{eqn:averaged_pdf} with $n = 15$.
-For both averaged and single cospectra, the observed histogram matches the theoretically expected distributions very closely.
-}}
-\label{fig:herx1_cs}
-\end{center}
-\end{figure*}
 
@@
 
-\begin{bf}
 Hercules X-1 (Her X-1) is a well-studied persistent X-ray binary pulsars \citep{tananbaum1972} in an X-ray binary comprising the neutron star itself and a stellar companion HZ Herculis \citep{davidsen1972,forman1972,bahcall1972} with a mass of $\sim$$2.2 \,\mathrm{M}_\odot$ \citep{reynolds1997,leahy2014} whose type varies between late-type A and early-type B with orbital phase \citep{anderson1994,cheng1995} .
@@
 While at low frequencies, the deviation from the expected noise distribution (a $\chi^2$ distribution centered on $2$) can be explained with the longer-term variability within the source, at frequencies above $5 \,\mathrm{Hz}$ or so, the periodogram displays the typical oscillatory pattern associated with dead time (see also \citealt{Bachetti+15}), suggesting that the standard distributions usually applied to periodograms will not produce unbiased results for these observations.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{nustar_cospectra.pdf}
+\caption{Upper panel: Cospectrum of a $10\,\mathrm{s}$ segment of the \nustar\ observations of Her X-1 (black).
+The cospectrum is created by observing the same source simultaneously, but statistically independently with the two identical FPMA and FPMB detectors.
+The rotational signal of the pulsar appears very clearly at $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$, but unlike the periodogram, the cospectrum is not affected by dead time.
+In red, we show the $p = 10^{-3}$ detection threshold, corrected for $6999$ trials (equivalent to the number of frequencies in the periodogram).
+Lower left:  histogram (grey) of the cospectral densities collected from 3260 individual cospectra between $50\,\mathrm{Hz}$ and $400\,\mathrm{Hz}$.
+In red, we also show the theoretically predicted Laplace distribution.
+Lower right: in black, the histogram of cospectral densities collected from $7$ averaged cospectra, each created from $15$ light curve segments of $200\,\mathrm{s}$ duration.
+Again, only cospectral densities between $50\,\mathrm{Hz}$ and $400\,\mathrm{Hz}$ were used in this figure.
+In red, we show the theoretical probability density function defined in Equation \ref{eqn:averaged_pdf} with $n = 15$.
+For both averaged and single cospectra, the observed histogram matches the theoretically expected distributions very closely.
+}
+\label{fig:herx1_cs}
+\end{center}
+\end{figure*}
 
@@
 This indicates that there is no evidence in the data that would argue against the observed cospectral densities being drawn from the distribution in Equation \ref{eqn:averaged_cdf}.
-\end{bf}
 
@@
 The cospectrum, on the other hand, only requires an estimate of the local variance in order to use the equations derived above, making it a far more convenient choice for periodicity detection.
-\textbf{By applying a straightforward correction using the differences between Fourier amplitudes derived from two independent detectors, cospectra can be corrected such that they follow the distributions derived here very closely.}
+By applying a straightforward correction using the differences between Fourier amplitudes derived from two independent detectors, cospectra can be corrected such that they follow the distributions derived here very closely.
 
-\textbf{We show that the same is true for observations of the bright neutron star X-ray binary Her X-1, where the intrinsic brightness of the source leads to strong modulations in the periodogram due to dead time.
+We show that the same is true for observations of the bright neutron star X-ray binary Her X-1, where the intrinsic brightness of the source leads to strong modulations in the periodogram due to dead time.
 For sources observed with \nustar\ and other instruments with a similar set-up of at least two redundant, identical detectors, one can take advantage of the independent, simultaneous observations in two detectors to form a cospectrum.
 While any (quasi-)periodic signal will appear weaker in the cospectrum than in the periodogram (even in the case, as presented here, where the signals in the two detectors are exactly in phase), it has the substantial advantage that it is less affected by dead time and other similar detector effects.
-We show that the cospectrum of the observations observed with \nustar\ follow the expected theoretical distributions very closely once the FAD correction is applied.}
+We show that the cospectrum of the observations observed with \nustar\ follow the expected theoretical distributions very closely once the FAD correction is applied.
 
